---
title: "Fixed effects and Random effects"
author: "Anna Werner & Susumu Shikano"
date: "`r Sys.Date()`"
output: html_document
---


```{r , include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# save ouputs
save.outputs <- 1
```

advantage:
more cases in here:
- 2 AfD cases are in here
- and the BSW cases are in here.

disadvantage:
- no year adjusted voting results, thus redistricting

## Load data

```{r}
load("Data/analysis_dat.RData")
```


## Switch for party selection and spill over

```{r}
party.selec.switch <- 1
spill.over.switch <- 0 # 1: with spill over
turnout.switch <- 0 # 1: turnout analysis
```

See below what value corresponds to what party selection.

```{r}
## Do not select only one party. This won't work!
## Party positions: 1=SPD, 2=CDU/CSU, 3=GRUENE, 4=FDP, 5=AfD, 6=DIE LINKE, 7=BSW

if (party.selec.switch==1) party.selection <- c(1:7)       # All parties: S C G F A L B
if (party.selec.switch==2) party.selection <- c(1:3)       # SPD, CDU/CSU, GRUENE (GRUENE in 2021+, AfD in 2025 only)
if (party.selec.switch==3) party.selection <- c(3:7)       # Without SPD and CDU/CSU (GRUENE pre-2021, AfD pre-2025)
if (party.selec.switch==4) party.selection <- c(1,3:7)     # Without CDU/CSU
if (party.selec.switch==5) party.selection <- c(2:7)       # Without SPD
if (party.selec.switch==6) party.selection <- c(1:3,5:7)   # Without FDP
if (party.selec.switch==7) party.selection <- c(1:5,7)     # Without DIE LINKE
if (party.selec.switch==8) party.selection <- c(1:2,4:7)   # Without GRUENE
if (party.selec.switch==9) party.selection <- c(1:4,6:7)   # Without AfD
if (party.selec.switch==10) party.selection <- c(1,3:7)    # Without CDU/CSU
if (party.selec.switch==11) party.selection <- c(2:4,6:7)  # Without SPD and DIE LINKE
if (party.selec.switch==12) party.selection <- c(1:6)      # Without BSW (traditional parties only)
```


```{r}
all.abs.votes <- all.abs.votes[,party.selection,,,]
all.treat <- all.treat[,party.selection,,,]

```


## Switch for denominator to compute the vote share

1: eligible citizens (for turnout in analysis, this option has to be selected.)
2: all voters
3: all valid votes 

```{r}

vote.denom <- 1

```

### Election results

We first compute the relative vote shares: 

```{r}

  all.rel.votes <-  all.abs.votes/all.abs.total[,rep(vote.denom,length(party.selection)),,,] 
  # If one uses valid votes: rep(3,6)
  # If one uses all voters: rep(2,6)
  # If one uses all eligible citizens: rep(1,6)

  # We delete the past results
  all.rel.votes <- all.rel.votes[,,,,1]

```



### Treatment

```{r}

all.treat.dummy <- all.treat
all.treat.dummy[all.treat.dummy>0] <- 1
all.treat.dummy <- all.treat.dummy[,,,,1]  # current-year treatment only; dim5=2 (previous-year) is for FD only

table(all.treat.dummy)



```

## index array

Year index

```{r}
year.index <- all.treat.dummy

year.index[, , , 1] <- 2025
year.index[, , , 2] <- 2021
year.index[, , , 3] <- 2017
year.index[, , , 4] <- 2013
year.index[, , , 5] <- 2009
year.index[, , , 6] <- 2005
year.index[, , , 7] <- 2002

table(year.index)

```

Party index

```{r}
party.index <- all.treat.dummy

party.labels <- c("SPD", "CDU", "GRU", "FDP", "AFD", "LIN", "BSW")[party.selection] # OLD: c("CDU","SPD","AFD","FDP","LIN","GRU")
for (i.party in 1:length(party.selection)){
  party.index[,i.party,,] <- party.labels[i.party]
}

table(party.index)
```

Modification of Wahlkreis index array

```{r}
all.wkr.index <- all.wkr.index[,party.selection,,]

```


### Special treatment of Greens and AfD

```{r}
print(paste("Number of non-missings",sum(is.na(all.rel.votes))))
if (party.selec.switch==2) {
  # Keep GRUENE only 2021+
  all.rel.votes[party.index=="GRU" & year.index < 2021] <- NA
  # Keep AfD only in 2025
  all.rel.votes[party.index=="AFD" & year.index < 2025] <- NA
}
if (party.selec.switch==3) {
  # Keep GRUENE only pre-2021 (2017 and earlier)
  all.rel.votes[party.index=="GRU" & year.index >= 2021] <- NA
  # Keep AfD only pre-2025 years
  all.rel.votes[party.index=="AFD" & year.index >= 2025] <- NA
}

print(paste("Number of non-missings after making Baerbock as KK",sum(is.na(all.rel.votes))))


```


# Fixed effect models

```{r}
library(lmtest)
library(sandwich)

for (i.smdpr in 1:2){ # first/second vote

  all.results <- vector("list",6) # empty for results

  votes.in.array <- all.rel.votes[,,i.smdpr,1:6]
  votes <- c(all.rel.votes[,,i.smdpr,1:6] )
  last.votes <- c(all.rel.votes[,,i.smdpr,2:7] )  # NOTE: introduces Nickell bias O(1/T)≈14% with T=6; retained to absorb AR dynamics (residuals among treated show β=0.09, R²=0.21 correlation with fitted values without it)
  treat <- c(all.treat.dummy[,,i.smdpr,1:6] )
  year  <- as.factor(c(year.index[,,i.smdpr,1:6] ))
  year <- relevel(year,ref="2009")
  party <- factor(c(party.index[, , i.smdpr, 1:6]),
                  levels = c("SPD", "CDU", "GRU", "FDP", "AFD", "LIN", "BSW"))
  party <- relevel(party, ref = "AFD")
  district <- as.factor(c(all.wkr.index[,,i.smdpr,1:6]))
  cross.treat <- c(all.treat.dummy[,,c(2,1)[i.smdpr],1:6] )

  lm.dat <- as.data.frame(votes)
  lm.dat <- cbind(lm.dat,treat,cross.treat,last.votes,year,party,district)
  
  print(paste("Number of observations: ",nrow(na.omit(lm.dat))))
  
  print("Number of observations per party: ")
  print(table(na.omit(lm.dat)$party))
  
  
  ## Model with year and party fixed effects.
  print(c("Erststimme (H1a)","Zweitstimme  (H1b)")[i.smdpr])
  options(max.print=100)
  print(summary(lm.fe1 <- lm(votes ~ treat + last.votes +
                               party * year + district*party )))

  print(c("Robust standard error"))
  r.se.fe1 <- sqrt(diag(vcovHC(lm.fe1, type = 'HC')))
  print(coef.fe1 <- coeftest(lm.fe1, vcov = vcovHC(lm.fe1, type = 'HC')))

  all.results[[6]] <- lm.fe1
  
  
  ## Model with cross treatment (contamination)

  print(c("Contamination Erststimme (H2b)","Contamination Zweitstimme  (H2a)")[i.smdpr])
  print(summary(lm.fe2 <- lm(votes ~ treat + cross.treat + last.votes +
                               party * year + district * party)))
  
  print(c("Robust standard error"))
  r.se.fe2 <- sqrt(diag(vcovHC(lm.fe2, type = 'HC')))
  print(lm.fe2.robust <- coeftest(lm.fe2, vcov = vcovHC(lm.fe2, type = 'HC')))
  
  
  all.results[[1]] <- lm.fe2
  all.results[[2]] <- lm.fe2.robust
  
  all.results[[3]] <- lm.dat
  all.results[[4]] <- votes.in.array
  
  all.results[[5]] <- r.se.fe2
  
  if (i.smdpr==1) all.results.smd <- all.results
  if (i.smdpr==2) all.results.pr <- all.results
}

out_dir <- file.path("BTW_2025", "Outputs", "Fixed_Effects", "/")
if (save.outputs == 1) {
  save(
    all.results.smd,
    all.results.pr,
    file = paste0(
      out_dir,
      "party.selec",
      party.selec.switch,
      "_denom", vote.denom,
      c("", "_spillover")[spill.over.switch + 1],
      c("", "_turnout")[turnout.switch + 1],
      "_", format(Sys.Date(), "%Y-%m-%d"),
      ".RData"
    )
  )
}

options(max.print=100)
```

## Some diagnostic checks


Checking residuals

```{r}
par(mfrow = c(1, 2), mar = c(4, 4.5, 3, 2))

for (i.smdpr in 1:2) {
  if (i.smdpr == 1) {
    results.te <- all.results.smd[[1]]
    data.te <- all.results.smd[[3]]
    vote.label <- "FPTP"
  } else {
    results.te <- all.results.pr[[1]]
    data.te <- all.results.pr[[3]]
    vote.label <- "PR"
  }
  
  data.te <- data.te[!is.na(data.te$votes) & !is.na(data.te$last.votes), ]
  
  residuals <- residuals(results.te)
  
  boxplot(residuals ~ data.te$treat,
          main = paste("Residuals:", vote.label),
          xlab = "Treatment", ylab = "Residuals",
          col = c("#E8F4F8", "#B8D4E8"),
          border = "#2C5F7C", las = 1)
  
  abline(h = 0, lty = 2, col = "gray50")
  grid(nx = NA, ny = NULL, col = "gray90", lty = 1)
  
  means <- tapply(residuals, as.factor(data.te$treat), mean)
  print(paste(vote.label, "- Mean residuals by treatment:"))
  print(means)
}
```
## Probably needs UPDATE FROM HERE

### Plotting the actual votes, predicted votes and residuals for treated district

```{r}
predicted <- fitted(results.te)  # fitted values from last loop iteration (PR)
residuals <- residuals(results.te)

par(mfrow=c(1,2))

plot(data.te$votes[data.te$treat==1] ~ predicted[data.te$treat==1],
     xlab="Predicted",ylab="Actual")
abline(coef=c(0,1),lty=2)
labels <- data.te[data.te$treat==1, c("year","party","district")]

text(predicted[data.te$treat==1], data.te$votes[data.te$treat==1],
     rownames(data.te)[data.te$treat==1],
     pos=1, cex=0.7)

summary(lm.rsd <- lm(residuals[data.te$treat==1] ~ predicted[data.te$treat==1]))
plot(residuals[data.te$treat==1] ~ predicted[data.te$treat==1],
     xlab="Predicted", ylab="Residuals")
abline(lm.rsd)

labels <- data.te[data.te$treat==1, c("year","party","district")]

text(predicted[data.te$treat==1], residuals[data.te$treat==1],
     rownames(data.te)[data.te$treat==1],
     pos=1, cex=0.7)


labels

```


```{r}

wkr.fe <- coefficients(results.te)
wkr.fe <- wkr.fe[grep("district",names(wkr.fe))]

hist(wkr.fe)

wkr.fe[order(wkr.fe)][1:5]

wkr.fe[order(wkr.fe,decreasing=TRUE)][1:5]

```


### Diagnostics I: Serial correlated errors

We regress u_t on u_t-1. If the effect is significant, there are serially correlated errors.

Note: We first have to assign individual districts at t to those at t-1.

```{r}

for (i.smdpr in 1:2){
  if (i.smdpr==1){
    to.be.tested <- all.results.smd[[1]]  # Diagnostics for the model Erststimme
    temp.dat <- all.results.smd[[3]]
    temp.depv <- all.results.smd[[4]]

  }else{
    to.be.tested <- all.results.pr[[1]] # Diagnostics for the model Zweitstimme
    temp.dat <- all.results.pr[[3]]
    temp.depv <- all.results.pr[[4]]
  }
  

residuals <- residuals(to.be.tested)
residuals <- array(residuals,dim=dim(temp.depv))

simple.district.idx <- all.wkr.index[,1,1,]

# stacking residuals at t and t-1
## loop from BTW 2021 to 2009

current.resid <- past.resid <- residuals[,,1:4]
for (i.btw in 1:4){
  # create the key to match the current and past districts
  key <- match(simple.district.idx[,i.btw],simple.district.idx[,i.btw+1])
  past.resid[,,i.btw] <- residuals[key,,i.btw+1]
}

print(c("Erststimme","Zweitstimme")[i.smdpr])
print(
   summary(lm.out <- lm(c(current.resid) ~ c(past.resid)))
)
plot(c(current.resid) ~ c(past.resid),
     ylab="t",xlab="t-1")
abline(lm.out)

}

```


### Exogeneity

```{r}
# Wooldridge's test

options(max.print=100)
# District candidacy: Treatment for SMD
 
dep.var <- c(all.rel.votes[,,1,2:5])
lagged.dv <- c(all.rel.votes[,,1,3:6])  # NOTE: inconsistent with static FE if Nickell bias concern is primary; included for consistency with main model
treat <- c(all.treat.dummy[,,1,2:5])
treat.t1 <- c(all.treat.dummy[,,1,1:4])
cross.treat <- c(all.treat.dummy[,,2,2:5])
cross.treat.t1 <- c(all.treat.dummy[,,2,1:4])

party.fe <- factor(c(party.index[,,1,2:5]),
                   levels=c("SPD","CDU","GRU","FDP","AFD","LIN","BSW"))
wkr.fe <- as.factor(c(all.wkr.index[,,1,2:5]))

year.fe  <- as.factor(c(year.index[,,1,2:5]))

summary(lm.out <- lm(dep.var ~ lagged.dv +
             treat + treat.t1+
             cross.treat + cross.treat.t1+
             year.fe*party.fe + party.fe*wkr.fe))

 print(c("Robust standard error"))
  r.se <- sqrt(diag(vcovHC(lm.out, type = 'HC')))
  print(lm.out.robust <- coeftest(lm.out, vcov = vcovHC(lm.out, type = 'HC')))

all.results.smd <- vector("list",3)
all.results.smd[[1]] <- lm.out
all.results.smd[[2]] <- lm.out.robust
all.results.smd[[3]] <- c("empty")
all.results.smd[[4]] <- c("empty")
all.results.smd[[5]] <- r.se


# PR list candidacy: Treatment for PR

all.results <- vector("list",3)

dep.var <- c(all.rel.votes[,,2,2:5])
lagged.dv <- c(all.rel.votes[,,2,3:6])  # NOTE: see FPTP block above
treat <- c(all.treat.dummy[,,2,2:5])
treat.t1 <- c(all.treat.dummy[,,2,1:4])
cross.treat <- c(all.treat.dummy[,,1,2:5])
cross.treat.t1 <- c(all.treat.dummy[,,1,1:4])

party.fe <- factor(c(party.index[,,2,2:5]),
                   levels=c("SPD","CDU","GRU","FDP","AFD","LIN","BSW"))
wkr.fe <- as.factor(c(all.wkr.index[,,2,2:5]))

year.fe  <- as.factor(c(year.index[,,2,2:5]))

summary(lm.out <- lm(dep.var ~ lagged.dv +
             treat + treat.t1+
             cross.treat + cross.treat.t1+
             year.fe*party.fe + party.fe*wkr.fe))

print(c("Robust standard error"))
  r.se <- sqrt(diag(vcovHC(lm.out, type = 'HC')))
  print(lm.out.robust <- coeftest(lm.out, vcov = vcovHC(lm.out, type = 'HC')))

  
all.results.pr <- vector("list",3)
all.results.pr[[1]] <- lm.out
all.results.pr[[2]] <- lm.out.robust
all.results.pr[[3]] <- c("empty")
all.results.pr[[4]] <- c("empty")
all.results.pr[[5]] <- r.se


save(all.results.smd,
     all.results.pr,
 file=(paste0(out_dir, "party.selec",
              party.selec.switch,
              "_denom",vote.denom,
              c("", "_spillover")[spill.over.switch + 1],
              c("", "_turnout")[turnout.switch + 1],
              "_exgnt_",
              format(Sys.Date(), "%Y-%m-%d"),
              ".RData")))

options(max.print=100)

```



