---
title: "First Differences"
author: "Anna Werner & Susumu Shikano"
date: "`r Sys.Date()`"
output: html_document
---

```{r , include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Load data

```{r}
load("Data/analysis_dat.RData")
```


## Switch for party selection

```{r}

## Do not select only one party. This won't work!

party.selection <- c(1:6) # C S A F L G
#party.selection <- c(3:6) # ohne C und S
#party.selection <- c(1:2) # nur C und S


```


```{r}
all.abs.votes <- all.abs.votes[,party.selection,,,]
all.treat     <- all.treat[,party.selection,,]

```


## Creating first difference

### Election results

We first compute the relative vote shares: 

```{r}

  all.rel.votes <-  all.abs.votes/all.abs.total[,rep(3,length(party.selection)),,,] 
  # If one uses valid votes: rep(3,length(party.selection))
  # If one uses all voters: rep(2,length(party.selection))
  # If one uses all eligible citizens: rep(1,length(party.selection))

```

Subsequently, we create the first differences:


```{r}
  all.rel.votes.fd <- all.rel.votes[,,,1:5,1] - all.rel.votes[,,,1:5,2]

  # Note that there will be no FD for 2002, therefore we have n(year)=5.

  hist(all.rel.votes.fd)
```

### Treatment

```{r}

all.treat.dummy <- all.treat
all.treat.dummy[all.treat.dummy>0] <- 1

table(all.treat.dummy)

all.treat.dummy.fd <- array(NA,dim=c(299,length(party.selection),2,5))
for (i.time in 1:5){
  all.treat.dummy.fd[,,,i.time] <-  
    all.treat.dummy[,,,i.time] - all.treat.dummy[,,,i.time+1] 
}

table(all.treat.dummy.fd)

```

## index array

Year index

```{r}
year.index <- all.treat.dummy.fd

year.index[,,,1] <- 2021
year.index[,,,2] <- 2017
year.index[,,,3] <- 2013
year.index[,,,4] <- 2009
year.index[,,,5] <- 2005

table(year.index)

```

Party index

```{r}
party.index <- all.treat.dummy.fd

party.labels <- c("CDU","SPD","AFD","FDP","LIN","GRU")[party.selection]
for (i.party in 1:length(party.selection)){
  party.index[,i.party,,] <- party.labels[i.party]
}

table(party.index)
```

## First difference models

```{r}


for (i.smdpr in 1:2){ # Loop over Erst- and Zweitstimme

  all.results <- vector("list",3) # empty list to store the results

  votes.in.array <- all.rel.votes.fd[,,i.smdpr,]
  votes <- c(all.rel.votes.fd[,,i.smdpr,] )
  treat <- c(all.treat.dummy.fd[,,i.smdpr,] )
  year  <- as.factor(c(year.index[,,i.smdpr,] ))
  party <- as.factor(c(party.index[,,i.smdpr,] ))
  cross.treat <- c(all.treat.dummy.fd[,,c(2,1)[i.smdpr],] )

  lm.dat <- cbind(votes,treat,year,party,cross.treat)
  
  ## Model with year and party fixed effects.
  print(c("Erststimme (H1a)","Zweitstimme  (H1b)")[i.smdpr])
  print(summary(lm.out <- lm(votes~ treat + year * party)))

  all.results[[1]] <- lm.out
  
  ## Model with cross treatment (contamination)

  print(c("Contamination on Erststimme (H2b)","Contamination on Zweitstimme  (H2a)")[i.smdpr])
  print(summary(lm.out <- lm(votes~ treat + cross.treat + year * party)))
  
  all.results[[2]] <- lm.out
  
  all.results[[3]] <- lm.dat
  all.results[[4]] <- votes.in.array
  
  if (i.smdpr==1) all.results.smd <- all.results
  if (i.smdpr==2) all.results.pr <- all.results
}


```

### Diagnostics I: Serial correlated errors

We regress u_t on u_t-1. If the effect is significant, there are serially correlated errors.

```{r}

for (i.smdpd in 1:2){
  if (i.smdpd==1){
    to.be.tested <- all.results.smd[[1]]  # Diagnostics for the model Erststimme
  }else{
    to.be.tested <- all.results.pr[[1]] # Diagnostics for the model Zweitstimme
  }
  
temp.dat <- all.results.smd[[3]]
temp.depv <- all.results.smd[[4]]

residuals <- residuals(to.be.tested)
residuals <- array(residuals,dim=dim(temp.depv))

# exclude outliers
#residuals0 <- residuals
#residuals[residuals>.15] <- NA

print(c("Erststimme","Zweitstimme")[i.smdpd])
print(
   
   summary(lm.out <- lm(c(residuals[,,5:2]) ~ c(residuals[,,4:1])))
)
plot(c(residuals[,,5:2]) ~ c(residuals[,,4:1]),
     ylab="t",xlab="t-1")
abline(lm.out)

}

```


### Diagnostics II: Exogeneity of the treatment

We regress the treatment variable on x_t on y_t-1. If the effect is significant, the assumption of the strict exogeneity is violated.

```{r}

# District candidacy: Treatment for SMD
summary(lm(c(all.treat[,,1,]) ~ c(all.rel.votes[,,1,,2])))

# PR list candidacy: Treatment for PR
summary(lm(c(all.treat[,,2,]) ~ c(all.rel.votes[,,2,,2])))


```


## Implementation of Chaisemartin and D’Haultfoeuille's estimator

Not completed...

```{r}

for (i.smdpr in 1:2){ # Loop over Erst- and Zweitstimme

  votes <- all.rel.votes.fd[,,i.smdpr,] 
  treat.raw <- all.treat.dummy[,,i.smdpr,] 
  treat <- all.treat.dummy.fd[,,i.smdpr,] 
  year  <- year.index[,,i.smdpr,] 
  party <- party.index[,,i.smdpr,] 
  cross.treat <- all.treat.dummy.fd[,,c(2,1)[i.smdpr],] 

  treat.types <- treat

  for (i.year in 1:dim(votes)[3]){ # loop over years
    treat.types[,,i.year][treat.raw[,,i.year]==1 & treat.raw[,,i.year+1]==1] <- 2
  }
  table(treat.types)
  
  did.plus <- mean(votes[treat.types==1],na.rm=T) - mean(votes[treat.types==0],na.rm=T)
  did.minus <- mean(votes[treat.types==2],na.rm=T) - mean(votes[treat.types==-1],na.rm=T)
  
  did.est <- (did.plus * sum(treat.types==1) + did.minus * sum(treat.types==-1))/sum(treat.types==1|treat.types==-1)

  print(did.est)
}
```



