---
title: "First Differences"
author: "Anna Werner & Susumu Shikano"
date: "`r Sys.Date()`"
output: html_document
---

```{r , include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stargazer)
library(MASS)
```


## Load data

```{r}
load("Data/analysis_dat.RData")
```

```{r}
# save ouputs
save.outputs <- 0
```

## Switch for party selection and spill over
 
```{r}
party.selec.switch <- 1 # see below for options
spill.over.switch <- 0 # 1: with spill over
turnout.switch <- 0 # 1: turnout analysis
```

See below what value corresponds to what party selection.

```{r}
## Do not select only one party. This won't work!
## Party positions: 1=SPD, 2=CDU/CSU, 3=GRUENE, 4=FDP, 5=AfD, 6=DIE LINKE, 7=BSW

if (party.selec.switch==1) party.selection <- c(1:7)       # All parties: S C G F A L B
if (party.selec.switch==2) party.selection <- c(1:3,5)     # SPD, CDU/CSU, GRUENE, AfD (GRUENE in 2021+, AfD in 2025 only)
if (party.selec.switch==3) party.selection <- c(3:7)       # Without SPD and CDU/CSU (GRUENE pre-2021, AfD pre-2025)
if (party.selec.switch==4) party.selection <- c(1,3:7)     # Without CDU/CSU
if (party.selec.switch==5) party.selection <- c(2:7)       # Without SPD
if (party.selec.switch==6) party.selection <- c(1:3,5:7)   # Without FDP
if (party.selec.switch==7) party.selection <- c(1:5,7)     # Without DIE LINKE
if (party.selec.switch==8) party.selection <- c(1:2,4:7)   # Without GRUENE
if (party.selec.switch==9) party.selection <- c(1:4,6:7)   # Without AfD
if (party.selec.switch==10) party.selection <- c(1,3:7)    # Without CDU/CSU
if (party.selec.switch==11) party.selection <- c(2:4,6:7)  # Without SPD and DIE LINKE
if (party.selec.switch==12) party.selection <- c(1:6)      # Without BSW (traditional parties only)
```

Reading adjacent matrix

```{r}
if (spill.over.switch>0){
  load("Data/btw_all_shape.RData")
  all.nb.mat <- all.nb.mat[3:8]
}

```



## Switch for denominator to compute the vote share

  1: eligible citizens (for turnout in analysis, this option has to be selected.)
  2: all voters
  3: all valid votes 

```{r}

vote.denom <- 1
if (turnout.switch==1) vote.denom <- 1 

```


all.abs.votes has 5 dimensions:
* **Constituencies** (1 … 299)
* **Parties** (1 … 7): S C G F A L B
* **First (FPTP) or second (PR) vote** (1, 2)
* **Year** (1 … 7): 2025, 2021, 2017, 2013, 2009, 2005, 2002
* **Current or previous results** (1, 2)

all.total has 5 dimensions:
* **Constituencies** (1 … 299)
* **Eligible voters, turnout, or valid votes** (1 … 3)
* **First or second vote** (1, 2)
* **Year** (1 … 7): 2025, 2021, 2017, 2013, 2009, 2005, 2002
* **Current or previous results** (1, 2)

all.treat has 5 dimensions:
* **Constituencies** (1 … 299)
* **Parties** (1 … 7): S C G F A L B
* **First or second vote** (1, 2)
* **Year** (1 … 7): 2025, 2021, 2017, 2013, 2009, 2005, 2002
* **Current or previous results** (1, 2)

```{r}
if (turnout.switch == 0){
  # Normal case: analyzing vote shares by party
  all.abs.votes <- all.abs.votes[, party.selection, , , ]
  all.treat     <- all.treat[, party.selection, , , ]
  all.rel.votes <- all.abs.votes /
    all.abs.total[, rep(vote.denom, length(party.selection)), , , , drop=FALSE]
} else {
  # Turnout analysis: treating turnout as the dependent variable
  all.abs.votes <- all.abs.total[, 2, , , , drop=FALSE]  # Waehlende (turnout)
  all.treat     <- all.treat[, party.selection, , , ]

  # Collapse treatment: if any party leader ran, treatment = 1
  temp <- apply(all.treat, c(1, 3, 4, 5), max)
  all.treat <- array(temp, dim=c(299, 1, 2, 7, 2))

  all.rel.votes <- all.abs.votes /
    all.abs.total[, rep(vote.denom, 1), , , , drop=FALSE]
}

```


## Creating first difference

### Election results

We create the first differences:


```{r}
# First differences: current results minus previous results
# Dimensions: [constituencies, parties, vote_type, years, current_vs_previous]
# We can compute FD for years 1-6 (2025, 2021, 2017, 2013, 2009, 2005)
# No FD for year 7 (2002) because there's no "previous" data

# data_then is already aligned to the corresponding current-year districts
# in Data_Prep.Rmd, so row-wise subtraction is the intended operation.
all.rel.votes.fd <- all.rel.votes[, , , 1:6, 1, drop=FALSE] -
                    all.rel.votes[, , , 1:6, 2, drop=FALSE]

# Note that there will be no FD for 2002, therefore we have n(year) = 6
all.rel.votes.fd <- array(all.rel.votes.fd,
                          dim = dim(all.rel.votes.fd)[1:4])

hist(all.rel.votes.fd)

# Check dimensions
dim(all.rel.votes.fd)  # Should be: 299, n_parties, 2, 6
```

### Treatment

Treatment leading candidate or not

```{r}
# Make all values > 0 to 1 (convert treatment levels to binary indicator)
all.treat.dummy <- all.treat
all.treat.dummy[all.treat.dummy > 0] <- 1

# Incumbent treatment: only values 3 or 4 indicate incumbent
all.treat.incumb <- ifelse(all.treat == 3 | all.treat == 4, 1, 0)

# Check distributions
table(all.treat.dummy)
apply(all.treat.dummy, 3, table)
table(all.treat.incumb)
apply(all.treat.incumb, 3, table)

# Treatment FD: mirrors vote FD exactly.
# dim5=1 = current-year treatment (data_treat ordering)
# dim5=2 = previous-year treatment reordered to current-year rows (data_treat_then)
# Row-wise subtraction is correct because both dims share the same row alignment.
all.treat.dummy.fd <- all.treat.dummy[,,,1:6,1,drop=FALSE] -
                      all.treat.dummy[,,,1:6,2,drop=FALSE]
all.treat.dummy.fd <- array(all.treat.dummy.fd, dim = dim(all.treat.dummy.fd)[1:4])

all.treat.incumb.fd <- all.treat.incumb[,,,1:6,1,drop=FALSE] -
                       all.treat.incumb[,,,1:6,2,drop=FALSE]
all.treat.incumb.fd <- array(all.treat.incumb.fd, dim = dim(all.treat.incumb.fd)[1:4])

# Check first difference distributions
table(all.treat.dummy.fd)
apply(all.treat.dummy.fd, 3, table)
table(all.treat.incumb.fd)
apply(all.treat.incumb.fd, 3, table)
```

```{r}
#  Check 2002 CDU Zweitstimmen treatment 
#table(all.treat.dummy[, 2, 2, 7])  # CDU, Zweitstimmen, 2002

# Check 2005 CDU Zweitstimmen treatment  
#table(all.treat.dummy[, 2, 2, 6])  # CDU, Zweitstimmen, 2005
# Should show: 0s and 1s

# Check FD for CDU Zweitstimmen 2005-2002
#table(all.treat.dummy.fd[, 2, 2, 6])  # Should show -1, 0, 1
```

The resulting all.treat.dummy.fd has:
- 299 districts
- 7 parties
- 2 tiers
- 6 time points



Incorporating spill-over treatment

```{r}
if (spill.over.switch > 0){
  for (i.tier in 1:2){
    for (i.btw in 1:6){
      temp.treat <- all.treat.dummy.fd[, , i.tier, i.btw]
      temp.mat <- as.matrix(all.nb.mat[[7 - i.btw]]) # 7-1=6: 2025
      diag(temp.mat) <- 1  # replace the diagonal elements with 1
      spill.treat <- temp.mat %*% temp.treat
      spill.treat[spill.treat > 0] <- 1
      spill.treat[spill.treat < 0] <- -1
      all.treat.dummy.fd[, , i.tier, i.btw] <- spill.treat
    }
  }
}

apply(all.treat.dummy.fd, 3, table)
```




## index array

Year index

```{r}
year.index <- all.treat.dummy.fd
year.index[, , , 1] <- 2025
year.index[, , , 2] <- 2021
year.index[, , , 3] <- 2017
year.index[, , , 4] <- 2013
year.index[, , , 5] <- 2009
year.index[, , , 6] <- 2005
table(year.index)

```

Party index

```{r}
party.index <- all.treat.dummy.fd

if (turnout.switch == 0){
  # Party order in arrays: SPD, CDU/CSU, GRUENE, FDP, AfD, DIE LINKE, BSW
  party.labels <- c("SPD", "CDU", "GRU", "FDP", "AFD", "LIN", "BSW")[party.selection]
  
  for (i.party in seq_along(party.selection)){
    party.index[, i.party, , ] <- party.labels[i.party]
  }
} else {
  party.labels <- "Turnout"
  party.index[, 1, , ] <- party.labels
}

table(party.index)
```

### Special treatment of Greens and AfD across major/minor specifications

```{r}
if (party.selec.switch==2) {
  # Major-party specification: keep GRUENE only in 2021 and 2025
  all.rel.votes.fd[party.index=="GRU" & year.index < 2021] <- NA
  # Major-party specification: keep AfD only in 2025
  all.rel.votes.fd[party.index=="AFD" & year.index < 2025] <- NA
}
if (party.selec.switch==3) {
  # Minor-party specification: keep GRUENE only pre-2021 (2017 and earlier)
  all.rel.votes.fd[party.index=="GRU" & year.index >= 2021] <- NA
  # Minor-party specification: keep AfD only pre-2025 years
  all.rel.votes.fd[party.index=="AFD" & year.index >= 2025] <- NA
}

```



## Some data checks

Descriptives for the FD and treatments

```{r}
out_dir <- file.path("BTW_2025", "Outputs", "First_Differencing", "Figures")
years <- c(2025, 2021, 2017, 2013, 2009, 2005)

# filename
generate_filename <- function(vote_type) {
  paste0(
    "party.selec", party.selec.switch,
    "_denom", vote.denom,
    if (spill.over.switch) "_spillover" else "",
    if (turnout.switch) "_turnout" else "",
    "_FD_Boxplots_", vote_type, "-",
    format(Sys.Date(), "%Y-%m-%d"),
    ".pdf"
  )
}

# plotting function
plot_fd_boxplots <- function(vote_dim, vote_label) {
  par(mfrow = c(2, 3), 
      mar = c(4, 4.5, 3, 1),
      oma = c(0, 0, 2, 0))
  
  for (i_year in seq_along(years)) {
    temp_fd <- unlist(all.rel.votes.fd[, , vote_dim, i_year])
    temp_treat <- unlist(all.treat.dummy.fd[, , vote_dim, i_year])
    
    valid_idx <- !is.na(temp_fd)
    
    boxplot(temp_fd[valid_idx] ~ temp_treat[valid_idx],
            main = years[i_year],
            xlab = "Treatment",
            ylab = "First Difference of Vote Share",
            col = c("#E8F4F8", "#B8D4E8"),
            border = "#2C5F7C",
            notch = FALSE,
            outline = TRUE,
            boxwex = 0.6,
            las = 1,
            cex.axis = 0.9,
            cex.lab = 1.0,
            cex.main = 1.1,
            font.main = 1)
    
    abline(h = 0, lty = 2, col = "gray50", lwd = 0.8)
    grid(nx = NA, ny = NULL, col = "gray90", lty = 1, lwd = 0.5)
  }
  
  mtext(paste("First Differencing Analysis:", vote_label, "Vote"), 
        outer = TRUE, cex = 1.3, font = 2)
}

# plot
if (save.outputs == 1) {
  pdf(file.path(out_dir, generate_filename("FPTP")), 
      width = 12, height = 8)
  plot_fd_boxplots(vote_dim = 1, vote_label = "FPTP")
  dev.off()
  
  pdf(file.path(out_dir, generate_filename("PR")), 
      width = 12, height = 8)
  plot_fd_boxplots(vote_dim = 2, vote_label = "PR")
  dev.off()
} else {
  plot_fd_boxplots(vote_dim = 1, vote_label = "FPTP")
  plot_fd_boxplots(vote_dim = 2, vote_label = "PR")
}
```




## First difference models

### 1. Data preparation

```{r}
library(sandwich)
library(lmtest)

# Optionally: previously we used HC SEs. But those may not adequately deal with the autocorrelation.
# Why not clustered SEs by constituency?

# Create lists to store data and results for both vote types
lm.dat.list <- vector("list", 2)
names(lm.dat.list) <- c("FPTP", "PR")

for (i.smdpr in 1:2){  # first/second vote
  votes.in.array <- all.rel.votes.fd[, , i.smdpr, ]
  votes <- c(all.rel.votes.fd[, , i.smdpr, ])
  treat <- c(all.treat.dummy.fd[, , i.smdpr, ])
  treat.incumb <- c(all.treat.incumb.fd[, , i.smdpr, ])
  year  <- as.factor(c(year.index[, , i.smdpr, ]))
  year <- relevel(year, ref = "2009")  # Change to middle year?
  
  # Party order: SPD, CDU, GRUENE, FDP, AfD, DIE LINKE, BSW
  party <- factor(c(party.index[, , i.smdpr, ]),
                  levels = c("SPD", "CDU", "GRU", "FDP", "AFD", "LIN", "BSW", "Turnout"))
  party <- relevel(party, ref = "AFD")
  
  cross.treat <- c(all.treat.dummy.fd[, , c(2,1)[i.smdpr], ])
  cross.treat.incumb <- c(all.treat.incumb.fd[, , c(2,1)[i.smdpr], ])
  
  # Each of the 299 constituencies appears multiple times (once per party per time period)
  # Use actual array dim[2] so this works for both normal (n parties) and turnout (1) modes
  n_series <- dim(all.treat.dummy.fd)[2]
  constituency <- rep(1:299, times = n_series * 6)

  lm.dat <- cbind(votes, treat, cross.treat, treat.incumb, cross.treat.incumb)
  lm.dat <- as.data.frame(lm.dat)
  lm.dat$year <- year
  lm.dat$party <- party
  lm.dat$wkr.name <- c(sapply(1:6, function(d) rep(all.wkr.name[, d], times = n_series)))
  lm.dat$wkr.nr   <- rep(1:299, times = n_series * 6)
  lm.dat$land      <- c(sapply(1:6, function(d) rep(all.land[, d], times = n_series)))

  lm.dat <- na.omit(lm.dat)
  
  # Store in list for later use
  lm.dat.list[[i.smdpr]] <- lm.dat
}

cat("\nData preparation complete. lm.dat.list created with", length(lm.dat.list), "elements.\n")
```

### 2. Data checking and analysis

```{r}
for (i.smdpr in 1:2){  # first/second vote
  lm.dat <- lm.dat.list[[i.smdpr]]
  vote.type <- c("Erststimme (FPTP)", "Zweitstimmen (PR)")[i.smdpr]
  
  cat("\n=== Data Investigation for", vote.type, "===\n")
  
  # 1. Summary statistics
  cat("\nDimensions of lm.dat:", nrow(lm.dat), "observations,", ncol(lm.dat), "variables\n")
  cat("Missing values (treatment):", sum(is.na(lm.dat$treat)), "\n")
  cat("Missing values (votes):", sum(is.na(lm.dat$votes)), "\n")
  
  # 2. Observations per year, party, and tier
  cat("\n--- Observations by Year and Party ---\n")
  obs.by.year.party <- xtabs(~year + party, data = lm.dat)
  print(obs.by.year.party)
  cat("\nTotal observations by year:\n")
  print(rowSums(obs.by.year.party))
  
  # 3. Treatment variable counts
  cat("\n--- Treatment Variable Counts ---\n")
  cat("Distribution of treatment variable:\n")
  print(table(lm.dat$treat, useNA = "ifany"))
  cat("\nTreatment: +1 =", sum(lm.dat$treat == 1, na.rm = TRUE), "observations\n")
  cat("Treatment: -1 =", sum(lm.dat$treat == -1, na.rm = TRUE), "observations\n")
  cat("Treatment:  0 =", sum(lm.dat$treat == 0, na.rm = TRUE), "observations\n")
  
  # 4. Cross-tabulation: Treatment by Year
  cat("\n--- Treatment by Year ---\n")
  treat.by.year <- xtabs(~year + treat, data = lm.dat)
  print(treat.by.year)
  
  # 5. Cross-tabulation: Treatment by Party
  cat("\n--- Treatment by Party ---\n")
  treat.by.party <- xtabs(~party + treat, data = lm.dat)
  print(treat.by.party)
}
```

### 2a Listing up the treated districts

```{r}
state.labels <- c("Schleswig-Holstein", "Hamburg", "Niedersachsen", "Bremen",
                  "NRW", "Hessen", "Rheinland-Pfalz", "Baden-Württemberg",
                  "Bayern", "Saarland", "Berlin", "Brandenburg",
                  "Mecklenburg-Vorpommern", "Sachsen", "Sachsen-Anhalt", "Thüringen")

# --- FPTP: one row per treated WK ---
cat("=== Vote FD for Treated WKs: FPTP (Erststimme) ===\n")
fptp <- lm.dat.list[["FPTP"]]

for (sign in c(1, -1)) {
  cat("\n", if (sign == 1) "Switched ON  (treat = +1)" else "Switched OFF (treat = -1)", "\n")
  sub <- fptp[fptp$treat == sign, c("wkr.nr", "wkr.name", "party", "year", "votes")]
  names(sub)[5] <- "vote.fd"
  print(sub[order(sub$party, sub$year), ], row.names = FALSE)
}

# --- PR: mean vote FD per state ---
# PR treatment is assigned at the state (Land) level: all WKs in a state share the same
# treatment indicator. Averaging vote.fd across WKs within a state therefore gives the
# mean PR vote share change across that state's constituencies in the treated election pair.
# `land` (integer 1–16, Bundeswahlleiter state ordering) was added to lm.dat during
# construction via all.land[,d] — one state index per WK per year, values always 1–16
# for years 2025–2005. Year 7 (2002) contains Wahlberechtigte counts instead of state
# indices because result0202 has a different column layout than data_now and col 3 is
# not the state index there; irrelevant here since FD only uses years 1–6.
cat("\n\n=== Vote FD for Treated WKs: PR (Zweitstimme) — mean per state ===\n")
pr <- lm.dat.list[["PR"]]

for (sign in c(1, -1)) {
  cat("\n", if (sign == 1) "Switched ON  (treat = +1)" else "Switched OFF (treat = -1)", "\n")
  sub <- pr[pr$treat == sign, ]
  # aggregate: for each state × party × year cell, average vote.fd over the WKs in that state
  agg <- aggregate(votes ~ land + party + year, data = sub, FUN = mean)
  agg$state <- state.labels[agg$land]
  agg <- agg[order(agg$party, agg$year, agg$land), c("state", "party", "year", "votes")]
  names(agg)[4] <- "mean.vote.fd"
  print(agg, row.names = FALSE)
}
```


### 3. Estimation

```{r}
for (i.smdpr in 1:2){  # first/second vote
  lm.dat <- lm.dat.list[[i.smdpr]]
  vote.type <- c("Erststimme (FPTP)", "Zweitstimmen (PR)")[i.smdpr]
  
  cat("\n\n=== Estimation for", vote.type, "===\n")
  all.results <- vector("list", 5)  # list for results
  votes.in.array <- all.rel.votes.fd[, , i.smdpr, ]
  
  print(c("DV: Erststimme", "DV: Zweitstimme")[i.smdpr])
  if (turnout.switch == 0){
    print(summary(lm.out <- lm(votes ~ treat +
                                  #treat.incumb + # control for Incumbency
                                  cross.treat +
                                  #cross.treat.incumb +
                                  year + year:party 
                                  , data = lm.dat))) 
  } else {
    print(summary(lm.out <- lm(votes ~ treat +
                                  #treat.incumb + # control for Incumbency
                                  cross.treat +
                                  #cross.treat.incumb +
                                  year 
                                  , data = lm.dat))) 
  }
  
  # HC option
  print(c("Robust standard error"))
  r.se <- sqrt(diag(vcovHC(lm.out, type = 'HC')))
  print(lm.out.robust <- coeftest(lm.out, vcov = vcovHC(lm.out, type = 'HC')))
  
  # OPTION: Cluster-robust standard errors
  #print("\n Cluster-Robust SEs (by Constituency)\n")
  #print("Accounts for serial correlation and heteroskedasticity within constituencies\n")
  #vcov_cluster <- vcovCL(lm.out, cluster = lm.dat$constituency, type = "HC1")
  #lm.out.cluster <- coeftest(lm.out, vcov = vcov_cluster)
  #print(lm.out.cluster)
  #r.se.cluster <- sqrt(diag(vcov_cluster))
  
  all.results[[1]] <- lm.out
  all.results[[2]] <- lm.out.robust
  #all.results[[2]] <- lm.out.cluster
  all.results[[3]] <- lm.dat
  all.results[[4]] <- votes.in.array
  all.results[[5]] <- r.se
  #all.results[[5]] <- r.se.cluster
  
  if (i.smdpr == 1) all.results.smd <- all.results
  if (i.smdpr == 2) all.results.pr <- all.results
}
```


```{r}
out_dir <- file.path("BTW_2025", "Outputs", "First_Differencing", "/")
if (save.outputs == 1) {
  save(
    all.results.smd,
    all.results.pr,
    file = paste0(
      out_dir,
      "party.selec",
      party.selec.switch,
      "_denom", vote.denom,
      c("", "_spillover")[spill.over.switch + 1],
      c("", "_turnout")[turnout.switch + 1],
      "_", format(Sys.Date(), "%Y-%m-%d"),
      ".RData"
    )
  )
}
```


```{r}
# Residuals vs Fitted
plot(lm.out, which = 1)

# Scale-Location plot
plot(lm.out, which = 3)

# Studentized residuals
plot(studres(lm.out) ~ fitted(lm.out),
     xlab = "Fitted Values",
     ylab = "Studentized Residuals",
     main = "Residuals vs Fitted")
abline(h = 0, lty = 2)
```

```{r }
## Diagnostic Tests
#print("Diagnostic Tests")


#for (i.smdpr in 1:2){
#  if (i.smdpr == 1){
#    lm.out <- all.results.smd[[1]]
#    temp.dat <- all.results.smd[[3]]
#  } else {
#   lm.out <- all.results.pr[[1]]
#    temp.dat <- all.results.pr[[3]]
#  }
  
#  print("")
#  print(c("ERSTSTIMME", "ZWEITSTIMME")[i.smdpr])
#  print("")
  
  # Breusch-Pagan Test
#  print("Breusch-Pagan Test (Heteroskedasticity):")
#  bp_test <- bptest(lm.out)
#  print(bp_test)
  
  # Visual diagnostic
#  par(mfrow = c(2, 2))
#  plot(lm.out)
#  par(mfrow = c(1, 1))
  
  # Shapiro-Wilk (normality)
#  print("Shapiro-Wilk Test (Normality of Residuals):")
#  shapiro.test(sample(residuals(lm.out), 5000))  # if n > 5000
#}
```



```{r}
if (turnout.switch==1|spill.over.switch==1) knitr::knit_exit()
```

### Diagnostics I: Serial correlated errors

We regress u_t on u_t-1. If the effect is significant, there are serially correlated errors.

Note: We first have to assign individual districts at t to those at t-1.

```{r}
for (i.smdpr in 1:2){
  if (i.smdpr == 1){
    to.be.tested <- all.results.smd[[1]]  # Diagnostics for the model Erststimme
    temp.dat <- all.results.smd[[3]]
    temp.depv <- all.results.smd[[4]]
  } else {
    to.be.tested <- all.results.pr[[1]]  # Diagnostics for the model Zweitstimme
    temp.dat <- all.results.pr[[3]]
    temp.depv <- all.results.pr[[4]]
  }
  
  residuals <- residuals(to.be.tested)
  if (party.selec.switch != 0){
    residuals <- array(residuals, dim = dim(temp.depv))
  } else {
    residuals <- array(residuals,
                       dim = c(dim(temp.depv)[1], 1, dim(temp.depv)[2]))
  }
  # Ensure residuals is always 3D — turnout mode drops the singleton party
  # dimension when indexing temp.depv, yielding a 2D [299, 6] matrix.
  if (length(dim(residuals)) == 2) {
    residuals <- array(residuals, dim = c(dim(residuals)[1], 1, dim(residuals)[2]))
  }

  # Stacking residuals at t and t-1
  ## Loop from BTW 2025 to 2005 (6 time periods, 5 pairs)
  current.resid <- past.resid <- residuals[, , 1:5, drop = FALSE]  # Changed from 1:4 to 1:5

  for (i.btw in 1:5){  # Changed from 1:4 to 1:5
    # Match year i.btw rows to year i.btw+1 rows using eligible voters.
    # data_then[curr_rows] carries the previous-year data in current-year row order,
    # data_now[prev_rows] carries the previous year in its own order — matching on
    # eligible voters identifies the same territory across both orderings.
    curr_rows <- (1 + 299*(i.btw-1)):(299*i.btw)
    prev_rows <- (1 + 299*i.btw):(299*(i.btw+1))
    key <- match(paste(data_then[curr_rows, 4], data_then[curr_rows, 6], sep = "_"),
                 paste(data_now[prev_rows,  4], data_now[prev_rows,  6], sep = "_"))
    past.resid[, , i.btw] <- residuals[key, , i.btw + 1]
  }
  
  print(c("Erststimme", "Zweitstimme")[i.smdpr])
  print(
    summary(lm.out <- lm(c(current.resid) ~ c(past.resid)))
  )
  
  plot(c(current.resid) ~ c(past.resid),
       ylab = "t", xlab = "t-1",
       main = paste(c("Erststimme", "Zweitstimme")[i.smdpr], 
                    "- Serial Correlation Test"))
  abline(lm.out)
}
```
Consider: Does the negative serial correlation make theoretical sense? It might indicate:
- Mean reversion in vote shares
- Voters "correcting" previous swings
- Backlash effects


### Diagnostics II: Exogeneity of the treatment

We are testing by using Wooldridge's test. That is, we include the treatment variable at t+1. For this purpose, we lose the most recent BTW 2021.

Also here, we have to take care about the redistricting.

```{r}
for (i.smdpr in 1:2){  # Loop over Erst- and Zweitstimme

  all.results.t1 <- vector("list", 5)  # empty list to store the results (changed from 3 to 5)

  votes.in.array <- all.rel.votes.fd[, , i.smdpr, 2:6]  # Changed from 2:5 to 2:6
  votes <- c(votes.in.array)
  treat <- c(all.treat.dummy.fd[, , i.smdpr, 2:6])  # Changed from 2:5 to 2:6
  treat.incumb <- c(all.treat.incumb.fd[, , i.smdpr, 2:6])  # Changed from 2:5 to 2:6
  year  <- as.factor(c(year.index[, , i.smdpr, 2:6]))  # Changed from 2:5 to 2:6
  year <- relevel(year, ref = "2009")  # Changed from "2005" to "2009"
  
  # Party order: SPD, CDU, GRUENE, FDP, AfD, DIE LINKE, BSW
  party <- factor(c(party.index[, , i.smdpr, 2:6]),  # Changed from 2:5 to 2:6
                  levels = c("SPD", "CDU", "GRU", "FDP", "AFD", "LIN", "BSW"))
  party <- relevel(party, ref = "AFD")
  
  cross.treat <- c(all.treat.dummy.fd[, , c(2,1)[i.smdpr], 2:6])  # Changed from 2:5 to 2:6
  cross.treat.incumb <- c(all.treat.incumb.fd[, , c(2,1)[i.smdpr], 2:6])  # Changed from 2:5 to 2:6
  
  # Treatment at t+1
  treat.t1 <- all.treat.dummy.fd[, , i.smdpr, 1:5, drop = FALSE]  # Changed from 1:4 to 1:5
  treat.t1 <- array(treat.t1, dim = dim(treat.t1)[c(1,2,4)])
  treat.incumb.t1 <- all.treat.incumb.fd[, , i.smdpr, 1:5, drop = FALSE]  # Changed from 1:4 to 1:5
  treat.incumb.t1 <- array(treat.incumb.t1, dim = dim(treat.incumb.t1)[c(1,2,4)])
  cross.treat.t1 <- all.treat.dummy.fd[, , c(2,1)[i.smdpr], 1:5, drop = FALSE]  # Changed from 1:4 to 1:5
  cross.treat.t1 <- array(cross.treat.t1, dim = dim(cross.treat.t1)[c(1,2,4)])
  cross.treat.incumb.t1 <- all.treat.incumb.fd[, , c(2,1)[i.smdpr], 1:5, drop = FALSE]  # Changed from 1:4 to 1:5
  cross.treat.incumb.t1 <- array(cross.treat.incumb.t1, dim = dim(cross.treat.incumb.t1)[c(1,2,4)])
  
  # Reorder t+1 treatment into year-t row positions using canonical territory IDs
  # (same source as data_treat_then in Data_Prep.Rmd, forward direction).
  # For each year i.btw, find which row in year i.btw-1 (t+1) holds the same
  # territory as each row in year i.btw (t). Hybrid: canonical match where
  # available, same-row fallback for boundary-change territories (NA key).
  canon.idx <- all.wkr.index[, 1, 1, ]  # [299, 7]: canonical IDs, year index 1=2025..7=2002
  for (i.btw in 2:6){  # Changed from 2:5 to 2:6 (loop over btw2021 to btw2005)
    key     <- match(canon.idx[, i.btw], canon.idx[, i.btw - 1])
    matched <- !is.na(key)
    fallback <- seq_len(299)

    reorder <- function(arr) {
      out <- arr
      if (is.null(dim(arr))) {
        # vector: turnout mode drops the singleton party dimension
        out[matched]  <- arr[key[matched]]
        out[!matched] <- arr[fallback[!matched]]
      } else {
        out[matched, ]  <- arr[key[matched], ]
        out[!matched, ] <- arr[fallback[!matched], ]
      }
      out
    }

    treat.t1[, , i.btw - 1]            <- reorder(treat.t1[, , i.btw - 1])
    treat.incumb.t1[, , i.btw - 1]     <- reorder(treat.incumb.t1[, , i.btw - 1])
    cross.treat.t1[, , i.btw - 1]      <- reorder(cross.treat.t1[, , i.btw - 1])
    cross.treat.incumb.t1[, , i.btw - 1] <- reorder(cross.treat.incumb.t1[, , i.btw - 1])
  }
  treat.t1 <- c(treat.t1)
  cross.treat.t1 <- c(cross.treat.t1)
  treat.incumb.t1 <- c(treat.incumb.t1)
  cross.treat.incumb.t1 <- c(cross.treat.incumb.t1)

  lm.dat <- cbind(votes, treat, year, party,
                  cross.treat, treat.incumb, cross.treat.incumb,
                  treat.t1, cross.treat.t1, treat.incumb.t1, cross.treat.incumb.t1)
  lm.dat <- as.data.frame(lm.dat)
  
  ## Model with cross treatment (contamination)

  print(c("DV: Erststimme", "DV: Zweitstimme")[i.smdpr])
  if (party.selec.switch != 0 && turnout.switch == 0){
    print(summary(lm.out <- lm(votes ~ treat + treat.t1 +
                                  #treat.incumb +
                                  #treat.incumb.t1 + # control for Incumbency
                                  cross.treat + cross.treat.t1 +
                                  #cross.treat.incumb + cross.treat.incumb.t1 +
                                  as.factor(year) + as.factor(year):as.factor(party),
                               data = lm.dat)))
  } else {
    print(summary(lm.out <- lm(votes ~ treat + treat.t1 +
                                  #treat.incumb +
                                  #treat.incumb.t1 + # control for Incumbency
                                  cross.treat + cross.treat.t1 +
                                  #cross.treat.incumb + cross.treat.incumb.t1 +
                                  as.factor(year),
                               data = lm.dat)))
  }
  
  print(c("Robust standard error"))
  r.se <- sqrt(diag(vcovHC(lm.out, type = 'HC')))
  print(lm.out.robust <- coeftest(lm.out, vcov = vcovHC(lm.out, type = 'HC')))
  
  print(c("Treatments Erststimmen", "Treatments Zweitstimmen")[i.smdpr])     
  print(table(treat))
  print(table(treat.t1))
  print(table(cross.treat))
  print(table(cross.treat.t1))

  all.results.t1[[1]] <- lm.out
  all.results.t1[[2]] <- lm.out.robust
  all.results.t1[[3]] <- lm.dat
  all.results.t1[[4]] <- votes.in.array
  all.results.t1[[5]] <- r.se
  
  if (i.smdpr == 1) all.results.t1.smd <- all.results.t1
  if (i.smdpr == 2) all.results.t1.pr <- all.results.t1
}

if (save.outputs == 1) {
  save(all.results.t1.smd,
     all.results.t1.pr,
     file = paste0(out_dir, "party.selec",
                   party.selec.switch,
                   "_denom", vote.denom,
                   c("", "_spillover")[spill.over.switch + 1],
                   c("", "_turnout")[turnout.switch + 1],
                   "_exgnt_",
                   format(Sys.Date(), "%Y-%m-%d"),
                   ".RData"))
  }

```
