---
title: "First Differences"
author: "Anna Werner & Susumu Shikano"
date: "`r Sys.Date()`"
output: html_document
---

```{r , include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stargazer)
```


## Load data

```{r}
load("Data/analysis_dat.RData")
```


## Switch for party selection and spill over
 
```{r}
party.selec.switch <-1
spill.over.switch <-1 # 1: with spill over
```

See below what value corresponds to what party selection.

```{r}

## Do not select only one party. This won't work!

if (party.selec.switch==0) party.selection <- c(1)        #  turnout
if (party.selec.switch==1) party.selection <- c(1:6)      # C S A F L G
if (party.selec.switch==2) party.selection <- c(1:2)      # nur C und S
if (party.selec.switch==3) party.selection <- c(3:6)      # ohne C und S
if (party.selec.switch==4) party.selection <- c(2:6)      # ohne C
if (party.selec.switch==5) party.selection <- c(1,3:6)    # ohne S
if (party.selec.switch==6) party.selection <- c(1:3,5:6) # ohne F
if (party.selec.switch==7) party.selection <- c(1:4,6)    # ohne Linke
if (party.selec.switch==8) party.selection <- c(1:5)     # ohne G
if (party.selec.switch==9) party.selection <- c(1:2,4:6)  # ohne AfD
if (party.selec.switch==10) party.selection <- c(2,4:6)    # ohne C und A
if (party.selec.switch==11) party.selection <- c(3:4,6)    # ohne C, S und L
```

Reading adjacent matrix

```{r}
if (spill.over.switch>0){
  load("Data/btw_all_shape.RData")
}

all.nb.mat <- all.nb.mat[3:7]
```




## Switch for denominator to compute the vote share

  1: eligible citizens (for turnout in analysis, this option has to be selected.)
  2: all voters
  3: all valid votes 

```{r}

vote.denom <- 1
if (party.selec.switch==0) vote.denom <- 1 

```


all.abs.votes has 5 dimensions:

- 299 districts
- 6 parties
- tier: FPTP or PR
- 6 elections
- current or last elections

all.treat has 4 dimensions:

- 299 districts
- 6 parties
- tier: FPTP or PR
- 6 elections

```{r}
if (party.selec.switch!=0){
  all.abs.votes <- all.abs.votes[,party.selection,,,]
  all.treat     <- all.treat[,party.selection,,]
}else{
  all.abs.votes <- all.abs.total[,rep(2,length(party.selection)),,,,drop=F] 
  temp <- apply(all.treat,c(1,3,4),max)
  all.treat <- array(temp,dim=c(299,1,2,6))    
}


all.rel.votes <-  all.abs.votes/
  all.abs.total[,rep(vote.denom,length(party.selection)),,,,drop=F] 

```


## Creating first difference

### Election results

We create the first differences:


```{r}
  all.rel.votes.fd <- all.rel.votes[,,,1:5,1,drop=F] - all.rel.votes[,,,1:5,2,drop=F]
  # Note that there will be no FD for 2002, therefore we have n(year)=5.
    all.rel.votes.fd <- array(all.rel.votes.fd,
                              dim=dim(all.rel.votes.fd)[1:4])

  hist(all.rel.votes.fd)
```

### Treatment

Treatment leading candidate or not

```{r}
# Make all values >0 to 1
all.treat.dummy <- all.treat
all.treat.dummy[all.treat.dummy>0] <- 1

all.treat.incumb <- ifelse(all.treat==3|all.treat==4,1,0)

table(all.treat.dummy)
apply(all.treat.dummy,3,table)
table(all.treat.incumb)
apply(all.treat.incumb,3,table)

all.treat.dummy.fd <- 
all.treat.incumb.fd <- 
  array(NA,dim=c(299,length(party.selection),2,5))

for (i.time in 1:5){ # loop over time
  all.treat.dummy.fd[,,,i.time] <-  
    all.treat.dummy[,,,i.time] - all.treat.dummy[,,,i.time+1] 

  all.treat.incumb.fd[,,,i.time] <-  
    all.treat.incumb[,,,i.time] - all.treat.incumb[,,,i.time+1] 

}

table(all.treat.dummy.fd)
table(all.treat.incumb.fd)
```

The resulting all.treat.dummy.fd has:
- 299 districts
- 6 parties
- 2 tiers
- 5 timepoints


Incomporting spill-over treatment

```{r}
if (spill.over.switch>0){
  
  for (i.btw in c(1:5)){
      temp.treat <- all.treat.dummy.fd[,,1,i.btw]
      temp.mat <- as.matrix(all.nb.mat[[6-i.btw]])
      diag(temp.mat) <- 1 # replace the diagonal elements with 1
      spill.treat <- temp.mat%*%temp.treat

      all.treat.dummy.fd[,,1,i.btw] <- spill.treat
  }
}

apply(all.treat.dummy.fd,3,table)

```




## index array

Year index

```{r}
year.index <- all.treat.dummy.fd

year.index[,,,1] <- 2021
year.index[,,,2] <- 2017
year.index[,,,3] <- 2013
year.index[,,,4] <- 2009
year.index[,,,5] <- 2005

table(year.index)

```

Party index

```{r}
party.index <- all.treat.dummy.fd

party.labels <- c("CDU","SPD","AFD","FDP","LIN","GRU")[party.selection]
if(party.selec.switch==0) party.labels <- "Turnout"
for (i.party in 1:length(party.selection)){
  party.index[,i.party,,] <- party.labels[i.party]
}

table(party.index)
```

## First difference models

```{r}
library(sandwich)
library(lmtest)

for (i.smdpr in 1:2){ # Loop over Erst- and Zweitstimme

  all.results <- vector("list",3) # empty list to store the results

  votes.in.array <- all.rel.votes.fd[,,i.smdpr,]
  votes <- c(all.rel.votes.fd[,,i.smdpr,] )
  treat <- c(all.treat.dummy.fd[,,i.smdpr,] )
  treat.incumb <- c(all.treat.incumb.fd[,,i.smdpr,] )
  year  <- as.factor(c(year.index[,,i.smdpr,] ))
  year <- relevel(year,ref="2005")
  party <- factor(c(party.index[,,i.smdpr,] ),
                  levels=c("CDU","SPD","AFD","FDP","LIN","GRU","Turnout")) # C S A F L G
  party <- relevel(party,ref="AFD")
  cross.treat <- c(all.treat.dummy.fd[,,c(2,1)[i.smdpr],] )
  cross.treat.incumb <- c(all.treat.incumb.fd[,,c(2,1)[i.smdpr],] )

  
  lm.dat <- cbind(votes,treat,cross.treat,treat.incumb,cross.treat.incumb)
  lm.dat <- as.data.frame(lm.dat)
  lm.dat$year <- year
  lm.dat$party <- party
  lm.dat <- na.omit(lm.dat)
  
  ## Model 

  print(c("DV: Erststimme","DV: Zweitstimme")[i.smdpr])
  if (party.selec.switch!=0){
      print(summary(lm.out <- lm(votes~ treat +
                                    #treat.incumb + # control for Incumbency
                                    cross.treat +
                                    #cross.treat.incumb +
                                    year + year:party 
                                    ,data=lm.dat))) 
  }else{
          print(summary(lm.out <- lm(votes~ treat +
                                    #treat.incumb + # control for Incumbency
                                    cross.treat +
                                    #cross.treat.incumb +
                                    year 
                                    ,data=lm.dat))) 
  }
  
  print(c("Robust standard error"))
  r.se <- sqrt(diag(vcovHC(lm.out, type = 'HC')))
  print(lm.out.robust <- coeftest(lm.out, vcov = vcovHC(lm.out, type = 'HC')))
  

  all.results[[1]] <- lm.out

  all.results[[2]] <- lm.out.robust
  
  all.results[[3]] <- lm.dat
  all.results[[4]] <- votes.in.array
  
  all.results[[5]] <- r.se
  
  if (i.smdpr==1) all.results.smd <- all.results
  if (i.smdpr==2) all.results.pr <- all.results
}

save(all.results.smd,
     all.results.pr,
  file=paste0("Data/Estimation_Results_PSelect",party.selec.switch,
             "_denom",vote.denom,c("","_spillover")[spill.over.switch+1],".RData"))

```




### Diagnostics I: Serial correlated errors

We regress u_t on u_t-1. If the effect is significant, there are serially correlated errors.

Note: We first have to assign individual districts at t to those at t-1.

```{r}

for (i.smdpr in 1:2){
  if (i.smdpr==1){
    to.be.tested <- all.results.smd[[1]]  # Diagnostics for the model Erststimme
    temp.dat <- all.results.smd[[3]]
    temp.depv <- all.results.smd[[4]]

  }else{
    to.be.tested <- all.results.pr[[1]]  # Diagnostics for the model Zweitstimme
    temp.dat <- all.results.pr[[3]]
    temp.depv <- all.results.pr[[4]]
  }
  

residuals <- residuals(to.be.tested)

if (party.selec.switch!=0){
  residuals <- array(residuals,dim=dim(temp.depv))
}else{
  residuals <- array(residuals,
                     dim=c(dim(temp.depv)[1],1,dim(temp.depv)[2]))
}


simple.district.idx <- all.wkr.index[,1,1,]

# stacking residuals at t and t-1
## loop from BTW 2021 to 2009
current.resid <- past.resid <- residuals[,,1:4,drop=F]

for (i.btw in 1:4){
  # create the key to match the current and past districts
  key <- match(simple.district.idx[,i.btw],simple.district.idx[,i.btw+1])
  past.resid[,,i.btw] <- residuals[key,,i.btw+1]
}

print(c("Erststimme","Zweitstimme")[i.smdpr])
print(
   summary(lm.out <- lm(c(current.resid) ~ c(past.resid)))
)
plot(c(current.resid) ~ c(past.resid),
     ylab="t",xlab="t-1")
abline(lm.out)

}

```



### Diagnostics II: Exogeneity of the treatment

We are testing by using Wooldridge's test. That is, we include the treatment variable at t+1. For this purpose, we lose the most recent BTW 2021.

Also here, we have to take care about the redistricting.


```{r}


for (i.smdpr in 1:2){ # Loop over Erst- and Zweitstimme

  all.results.t1 <- vector("list",3) # empty list to store the results

  votes.in.array <- all.rel.votes.fd[,,i.smdpr,2:5]
  votes <- c(votes.in.array)
  treat <- c(all.treat.dummy.fd[,,i.smdpr,2:5] )
  treat.incumb <- c(all.treat.incumb.fd[,,i.smdpr,2:5] )
  year  <- as.factor(c(year.index[,,i.smdpr,2:5] ))
  year <- relevel(year,ref="2005")
  party <- factor(c(party.index[,,i.smdpr,2:5] ),
                  levels=c("CDU","SPD","AFD","FDP","LIN","GRU")) # C S A F L G
  party <- relevel(party,ref="AFD")
  cross.treat <- c(all.treat.dummy.fd[,,c(2,1)[i.smdpr],2:5] )
  cross.treat.incumb <- c(all.treat.incumb.fd[,,c(2,1)[i.smdpr],2:5] )
  
  # treatment at t+1
  treat.t1 <- all.treat.dummy.fd[,,i.smdpr,1:4,drop=F]
  treat.t1 <- array(treat.t1,dim=dim(treat.t1)[c(1,2,4)])
  treat.incumb.t1 <- all.treat.incumb.fd[,,i.smdpr,1:4,drop=F] 
  treat.incumb.t1 <- array(treat.incumb.t1,dim=dim(treat.incumb.t1)[c(1,2,4)])
  cross.treat.t1 <- all.treat.dummy.fd[,,c(2,1)[i.smdpr],1:4,drop=F] 
  cross.treat.t1 <- array(cross.treat.t1,dim=dim(cross.treat.t1)[c(1,2,4)])
  cross.treat.incumb.t1 <- all.treat.incumb.fd[,,c(2,1)[i.smdpr],1:4,drop=F]
  cross.treat.incumb.t1 <- array(cross.treat.incumb.t1,dim=dim(cross.treat.incumb.t1)[c(1,2,4)])
  
  # we assign the correct district
  simple.district.idx <- all.wkr.index[,1,1,]
  for (i.btw in 2:5){ # loop over btw2017 to btw2005
    key <- match(simple.district.idx[,i.btw],simple.district.idx[,i.btw-1])
    treat.t1[,,i.btw-1] <- treat.t1[key,,i.btw-1]
    treat.incumb.t1[,,i.btw-1] <- treat.incumb.t1[key,,i.btw-1]
    cross.treat.t1[,,i.btw-1] <- cross.treat.t1[key,,i.btw-1]
    cross.treat.incumb.t1[,,i.btw-1] <- cross.treat.incumb.t1[key,,i.btw-1]
  }
  treat.t1 <- c(treat.t1)
  cross.treat.t1 <- c(cross.treat.t1)
  treat.incumb.t1 <- c(treat.incumb.t1)
  cross.treat.incumb.t1 <- c(cross.treat.incumb.t1)

  lm.dat <- cbind(votes,treat,year,party,
                  cross.treat,treat.incumb, cross.treat.incumb,
                  treat.t1,cross.treat.t1,treat.incumb.t1, cross.treat.incumb.t1)
  lm.dat <- as.data.frame(lm.dat)
  
  ## Model with cross treatment (contamination)

  print(c("DV: Erststimme","DV: Zweitstimme")[i.smdpr])
  if (party.selec.switch!=0){
      print(summary(lm.out <- lm(votes~ treat + treat.t1 +
                                    #treat.incumb + 
                                    #treat.incumb.t1 +# control for Incumbency
                                    cross.treat + cross.treat.t1 +
                                    #cross.treat.incumb + cross.treat.incumb.t1 +
                               as.factor(year) + as.factor(year):as.factor(party),
                             data=lm.dat)))
  }else{
      print(summary(lm.out <- lm(votes~ treat + treat.t1 +
                                    #treat.incumb + 
                                    #treat.incumb.t1 +# control for Incumbency
                                    cross.treat + cross.treat.t1 +
                                    #cross.treat.incumb + cross.treat.incumb.t1 +
                               as.factor(year) + as.factor(year),
                             data=lm.dat)))
  }
  
  print(c("Robust standard error"))
  r.se <- sqrt(diag(vcovHC(lm.out, type = 'HC')))
  print(lm.out.robust <- coeftest(lm.out, vcov = vcovHC(lm.out, type = 'HC')))
  
  print(c("Treatments Erststimmen","Treatments Zweitstimmen")[i.smdpr])     
        print(table(treat))
        print(table(treat.t1))
        print(table(cross.treat))
        print(table(cross.treat.t1))


  all.results.t1[[1]] <- lm.out

  all.results.t1[[2]] <- lm.out.robust
  
  all.results.t1[[3]] <- lm.dat
  all.results.t1[[4]] <- votes.in.array
  
  all.results.t1[[5]] <- r.se
  
  if (i.smdpr==1) all.results.t1.smd <- all.results.t1
  if (i.smdpr==2) all.results.t1.pr <- all.results.t1
}

save(all.results.t1.smd,
     all.results.t1.pr,
  file=paste0("Data/Exogeneity_Results_PSelect",party.selec.switch,
              "_denom",vote.denom,c("","_spillover")[spill.over.switch+1],".RData"))
```



