---
title: "Fixed effects and Random effects"
author: "Anna Werner & Susumu Shikano"
date: "`r Sys.Date()`"
output: html_document
---


```{r , include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Load data

```{r}
load("Data/analysis_dat.RData")
```



## Switch for party selection

```{r}

## Do not select only one party. This won't work!

party.selection <- c(1:6) # C S A F L G
#party.selection <- c(3:6) # ohne C und S
#party.selection <- c(1:2) # nur C und S
#party.selection <- c(1:4,6) # ohne Linke

```


```{r}
all.abs.votes <- all.abs.votes[,party.selection,,,]
all.treat     <- all.treat[,party.selection,,]

```



### Election results

We first compute the relative vote shares: 

```{r}

  all.rel.votes <-  all.abs.votes/all.abs.total[,rep(1,length(party.selection)),,,] 
  # If one uses valid votes: rep(3,6)
  # If one uses all voters: rep(2,6)
  # If one uses all eligible citizens: rep(1,6)

  # We delete the past results
  all.rel.votes <- all.rel.votes[,,,,1]

```



### Treatment

```{r}

all.treat.dummy <- all.treat
all.treat.dummy[all.treat.dummy>0] <- 1

table(all.treat.dummy)



```

## index array

Year index

```{r}
year.index <- all.treat.dummy

year.index[,,,1] <- 2021
year.index[,,,2] <- 2017
year.index[,,,3] <- 2013
year.index[,,,4] <- 2009
year.index[,,,5] <- 2005
year.index[,,,6] <- 2002

table(year.index)

```

Party index

```{r}
party.index <- all.treat.dummy

party.labels <- c("CDU","SPD","AFD","FDP","LIN","GRU")[party.selection]
for (i.party in 1:length(party.selection)){
  party.index[,i.party,,] <- party.labels[i.party]
}

table(party.index)
```

Modification of wahlkreis index array

```{r}
all.wkr.index <- all.wkr.index[,party.selection,,]

```

# Fixed effect models


```{r}

for (i.smdpr in 1:2){ # Loop over Erst- and Zweitstimme

  all.results <- vector("list",3) # empty list to store the results

  votes.in.array <- all.rel.votes[,,i.smdpr,1:5]
  votes <- c(all.rel.votes[,,i.smdpr,1:5] )
  last.votes <- c(all.rel.votes[,,i.smdpr,2:6] )
  treat <- c(all.treat.dummy[,,i.smdpr,1:5] )
  year  <- as.factor(c(year.index[,,i.smdpr,1:5] ))
  year <- relevel(year,ref="2005")
  party <- factor(c(party.index[,,i.smdpr,1:5] ),
                  levels=c("CDU","SPD","AFD","FDP","LIN","GRU")) # C S A F L G
  party <- relevel(party,ref="CDU")
  district <- as.factor(c(all.wkr.index[,,i.smdpr,1:5]))
  cross.treat <- c(all.treat.dummy[,,c(2,1)[i.smdpr],1:5] )

  lm.dat <- as.data.frame(votes)
  lm.dat <- cbind(lm.dat,treat,cross.treat,last.votes,year,party,district)
  
  ## Model with year and party fixed effects.
  print(c("Erststimme (H1a)","Zweitstimme  (H1b)")[i.smdpr])
  options(max.print=100)
  print(summary(lm.out <- lm(votes~ treat + last.votes+
                               year * party + district*party )))

  all.results[[1]] <- lm.out
  
  ## Model with cross treatment (contamination)

  print(c("Contamination Erststimme (H2b)","Contamination Zweitstimme  (H2a)")[i.smdpr])
  print(summary(lm.out <- lm(votes~ treat + cross.treat + 
                               last.votes + year * party + district*party)))
  
  all.results[[2]] <- lm.out
  
  all.results[[3]] <- lm.dat
  all.results[[4]] <- votes.in.array
  
  if (i.smdpr==1) all.results.smd <- all.results
  if (i.smdpr==2) all.results.pr <- all.results
}

options(max.print=10000)
```

## Some diagnostic checks


Checking residuals

```{r}

results.te <- all.results.smd[[1]] # nur für Erststimme? 

#plot(results.te)


data.te <- all.results.smd[[3]]

data.te <- data.te[!is.na(data.te$votes)&!is.na(data.te$last.votes),]
dim(data.te)


residuals <- residuals(results.te)
predicted <- predict(results.te)

length(residuals)


plot(residuals ~ data.te$treat)
tapply(residuals , as.factor( data.te$treat),mean)



```

### Plotting the actual votes, predicted votes and residuals for treated district

```{r}
par(mfrow=c(1,2)) 

plot(data.te$votes[data.te$treat==1] ~ predicted[data.te$treat==1],
     xlab="Predicted",ylab="Actual")
abline(coef=c(0,1),lty=2)
labels <- data.te[data.te$treat==1,3:5]

text( predicted[data.te$treat==1],data.te$votes[data.te$treat==1] ,
     rownames(data.te)[data.te$treat==1],
     pos=1,cex=0.7)

summary(lm.rsd <- lm(residuals[data.te$treat==1] ~ predicted[data.te$treat==1]))
plot(residuals[data.te$treat==1] ~ predicted[data.te$treat==1],
     xlab="Predicted",ylab="Rersiduals")
abline(lm.rsd)

labels <- data.te[data.te$treat==1,3:5]

text( predicted[data.te$treat==1],residuals[data.te$treat==1] ,
     rownames(data.te)[data.te$treat==1],
     pos=1,cex=0.7)


labels

```


```{r}

wkr.fe <- coefficients(results.te)
wkr.fe <- wkr.fe[grep("district",names(wkr.fe))]

hist(wkr.fe)

wkr.fe[order(wkr.fe)][1:5]

wkr.fe[order(wkr.fe,decreasing=TRUE)][1:5]

```

### Exogeneity

```{r}
# Wooldridge's test

options(max.print=100)
# District candidacy: Treatment for SMD

dep.var <- c(all.rel.votes[,,1,2:5])
lagged.dv <- c(all.rel.votes[,,1,3:6])
treat <- c(all.rel.votes[,,1,2:5])
treat.t1 <- c(all.rel.votes[,,1,1:4])
cross.treat <- c(all.rel.votes[,,2,2:5])
cross.treat.t1 <- c(all.rel.votes[,,2,1:4])

party.fe <- factor(c(party.index[,,1,2:5]),
                   levels=c("CDU","SPD","AFD","FDP","LIN","GRU"))
wkr.fe <- as.factor(c(all.wkr.index[,,1,2:5]))

year.fe  <- as.factor(c(year.index[,,1,1:4] ))

summary(lm(dep.var ~ lagged.dv +
             treat + treat.t1+
             cross.treat + cross.treat.t1+
             year.fe*party.fe + party.fe*wkr.fe))


# PR list candidacy: Treatment for PR

dep.var <- c(all.rel.votes[,,2,2:5])
lagged.dv <- c(all.rel.votes[,,2,3:6])
treat <- c(all.rel.votes[,,2,2:5])
treat.t1 <- c(all.rel.votes[,,2,1:4])
cross.treat <- c(all.rel.votes[,,1,2:5])
cross.treat.t1 <- c(all.rel.votes[,,1,1:4])

party.fe <- factor(c(party.index[,,2,2:5]),
                   levels=c("CDU","SPD","AFD","FDP","LIN","GRU"))
wkr.fe <- as.factor(c(all.wkr.index[,,2,2:5]))

year.fe  <- as.factor(c(year.index[,,2,1:4] ))

summary(lm(dep.var ~ lagged.dv +
             treat + treat.t1+
             cross.treat + cross.treat.t1+
             year.fe*party.fe + party.fe*wkr.fe))



options(max.print=10000)

```


### Diagnostics I: Serial correlated errors

We regress u_t on u_t-1. If the effect is significant, there are serially correlated errors.

Note: We first have to assign individual districts at t to those at t-1.

```{r}

for (i.smdpr in 1:2){
  if (i.smdpr==1){
    to.be.tested <- all.results.smd[[1]]  # Diagnostics for the model Erststimme
    temp.dat <- all.results.smd[[3]]
    temp.depv <- all.results.smd[[4]]

  }else{
    to.be.tested <- all.results.pr[[1]] # Diagnostics for the model Zweitstimme
    temp.dat <- all.results.pr[[3]]
    temp.depv <- all.results.pr[[4]]
  }
  

residuals <- residuals(to.be.tested)
residuals <- array(residuals,dim=dim(temp.depv))

simple.district.idx <- all.wkr.index[,1,1,]

# stacking residuals at t and t-1
## loop from BTW 2021 to 2009

current.resid <- past.resid <- residuals[,,1:4]
for (i.btw in 1:4){
  # create the key to match the current and past districts
  key <- match(simple.district.idx[,i.btw],simple.district.idx[,i.btw+1])
  past.resid[,,i.btw] <- residuals[key,,i.btw+1]
}

print(c("Erststimme","Zweitstimme")[i.smdpr])
print(
   summary(lm.out <- lm(c(current.resid) ~ c(past.resid)))
)
plot(c(current.resid) ~ c(past.resid),
     ylab="t",xlab="t-1")
abline(lm.out)

}

```

