---
title: "Analysis_FE"
author: "Anna Werner"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plm)
```

## Load Data

```{r}
load("Data/analysis_dat.RData")
```

## Relative vote share

```{r}
share_now <- data_now[,12:23]/data_now[,rep(c(10,11),6)]
share_then <- data_then[,12:23]/data_then[,rep(c(10,11),6)]

share_dif0 <- share_now - share_then

```


# Some descriptive observations


Here, we create an array:

1: districts
2: parties
3: district vs PR

```{r}

share_dif_arr <- array(NA,dim=c(nrow(share_dif0),ncol(share_dif0)/2,2))

share_dif_arr[,,1] <- as.matrix(share_dif0[,seq(1,11,by=2)])
share_dif_arr[,,2] <- as.matrix(share_dif0[,seq(2,12,by=2)])

treat_mat0 <- data_treat[,12:23]

treat_arr <- array(NA,dim=c(nrow(treat_mat0),ncol(treat_mat0)/2,2))

treat_arr[,,1] <- as.matrix(treat_mat0[,seq(1,11,by=2)])
treat_arr[,,2] <- as.matrix(treat_mat0[,seq(2,12,by=2)])

print("Frequencies of treatment values (district)")
table(treat_arr[,,1])
print("Frequencies of treatment values (PR)")
table(treat_arr[,,2])

par(mfrow=c(1,2))
for (i.fig in 1:2){
  boxplot(c(share_dif_arr[,,i.fig]) ~ c(treat_arr[,,i.fig]),
          xlab="Treatment",ylab="Vote difference in %-p",
          main=c("District cand","Party list")[i.fig])
  abline(h=0,lty=2)
}




```



# Specifying treatments and data selection

```{r}
 treat_mat0 <- data_treat[,12:23]
# treat_mat0[treat_mat0>0] <- 1    # All treatment included

#treat_mat0[treat_mat0==2|treat_mat0==4|treat_mat0==6] <- 0    # only chancellor candidates
#treat_mat0[treat_mat0>0] <- 1    

 treat_mat0[treat_mat0==1|treat_mat0==3|treat_mat0==5] <- 0    # only leading candidates without cc
 treat_mat0[treat_mat0>0] <- 1

# treat_mat0[treat_mat0==1|treat_mat0==2] <- 0    # candidates who have run for direct mandate before
# treat_mat0[treat_mat0>0] <- 1

```



# Data selection

```{r}
which.vote <- 2  # Erststimme or Zweitstimme?
contamination <- 0 # Contamination yes or no?


if (contamination==0){
  if (which.vote==1){
    share_dif <- share_dif0[,seq(1,11,by=2)]    # H1a: Direktmandat auf Erststimmenergebnis
    treat_mat <- treat_mat0[,seq(1,11,by=2)]
  } else {
    share_dif <- share_dif0[,seq(2,12,by=2)]    # H2a: Listenplatz auf Zweitstimmenergebnis
    treat_mat <- treat_mat0[,seq(2,12,by=2)]
  }                                             # H2?: Interaktion: Direktkandidatur * Parteiliste auf Zweitstimmen
} else {
  if (which.vote==1){
    share_dif <- share_dif0[,seq(1,11,by=2)]    # H2b: Liste auf Erststimmenergebnis
    treat_mat <- treat_mat0[,seq(2,12,by=2)]
  } else {
    share_dif <- share_dif0[,seq(2,12,by=2)]    # H1b: Direktmandat auf Zweitstimmenergebnis
    treat_mat <- treat_mat0[,seq(1,11,by=2)]
  }
}

```

# Year index

```{r}
year <- data_now$Jahr

year <- rep(year,ncol(share_dif))

```


# Party index

```{r}
party <- c("CDU/CSU", "SPD", "AfD", "FDP", "Linke", "Grüne")

party <- rep(party,each=nrow(share_dif))

```


## Regression

```{r}

summary(lm(unlist(share_dif)~unlist(treat_mat)  ))
summary(lm(unlist(share_dif)~unlist(treat_mat) * as.factor(party) * as.factor(year)) )

```

## Relative vote share - without CDU

```{r}
share_dif1 <- share_dif0[,3:12]  # Which parties? 
```

# Specifying treatments - without CDU

```{r}
treat_mat1 <- data_treat[,14:23]
treat_mat1[treat_mat1>0] <- 1    # All treatment included

# treat_mat1[treat_mat1==2|treat_mat1==4|treat_mat1==6] <- 0    # only chancellor candidates
# treat_mat1[treat_mat1>0] <- 1    

# treat_mat1[treat_mat1==1|treat_mat1==3|treat_mat1==5] <- 0    # only leading candidates without cc
# treat_mat1[treat_mat1>0] <- 1

# treat_mat1[treat_mat1==1|treat_mat1==2] <- 0    # candidates who have run for direct mandate before
# treat_mat1[treat_mat1>0] <- 1

```

# Data selection - without CDU

```{r}
which.vote <- 1   # Erststimme or Zweitstimme?
contamination <- 1 # Contamination yes or no?


if (contamination==0){
  if (which.vote==1){
    share_dif_saflg <- share_dif1[,seq(1,9,by=2)]    # H1a: Direktmandat auf Erststimmenergebnis
    treat_mat_saflg <- treat_mat1[,seq(1,9,by=2)]
  } else {
    share_dif_saflg <- share_dif1[,seq(2,10,by=2)]    # H2a: Listenplatz auf Zweitstimmenergebnis
    treat_mat_saflg <- treat_mat1[,seq(2,10,by=2)]
  }                                                  # H2?: Interaktion: Direktkandidatur * Parteiliste auf Zweitstimmen
} else {
  if (which.vote==1){
    share_dif_saflg <- share_dif1[,seq(1,9,by=2)]    # H2b: Liste auf Erststimmenergebnis
    treat_mat_saflg <- treat_mat1[,seq(2,10,by=2)]
  } else {
    share_dif_saflg <- share_dif1[,seq(2,10,by=2)]    # H1b: Direktmandat auf Zweitstimmenergebnis
    treat_mat_saflg <- treat_mat1[,seq(1,9,by=2)]
  }
}
```

# Year index - without CDU

```{r}
year1 <- data_now$Jahr
year1 <- rep(year1,ncol(share_dif_saflg))
```

# Party index - without CDU

```{r}
party_saflg <- c("SPD", "AfD", "FDP", "Linke", "Grüne")
party_saflg <- rep(party_saflg,each=nrow(share_dif_saflg))
```

## Regression - without CDU

```{r}
summary(lm(unlist(share_dif_saflg)~unlist(treat_mat_saflg)  ))
summary(lm(unlist(share_dif_saflg)~unlist(treat_mat_saflg) * as.factor(party_saflg) * as.factor(year1)) )
```


## Relative vote share - without SPD

```{r}
share_dif2 <- share_dif0[,c(1:2,5:12)]  # Which parties? 
```

# Specifying treatments - without SPD

```{r}
treat_mat2 <- data_treat[,c(12,13,16:23)]
treat_mat2[treat_mat2>0] <- 1    # All treatment included

# treat_mat2[treat_mat2==2|treat_mat2==4|treat_mat2==6] <- 0    # only chancellor candidates
# treat_mat2[treat_mat2>0] <- 1    

# treat_mat2[treat_mat2==1|treat_mat2==3|treat_mat2==5] <- 0    # only leading candidates without cc
# treat_mat2[treat_mat2>0] <- 1

# treat_mat2[treat_mat2==1|treat_mat2==2] <- 0    # candidates who have run for direct mandate before
# treat_mat2[treat_mat2>0] <- 1

```

# Data selection - without SPD

```{r}
which.vote <- 1    # Erststimme or Zweitstimme?
contamination <- 1 # Contamination yes or no?


if (contamination==0){
  if (which.vote==1){
    share_dif_caflg <- share_dif2[,seq(1,9,by=2)]    # H1a: Direktmandat auf Erststimmenergebnis
    treat_mat_caflg <- treat_mat2[,seq(1,9,by=2)]
  } else {
    share_dif_caflg <- share_dif2[,seq(2,10,by=2)]    # H2a: Listenplatz auf Zweitstimmenergebnis
    treat_mat_caflg <- treat_mat2[,seq(2,10,by=2)]
  }                                                  # H2?: Interaktion: Direktkandidatur * Parteiliste auf Zweitstimmen
} else {
  if (which.vote==1){
    share_dif_caflg <- share_dif2[,seq(1,9,by=2)]    # H2b: Liste auf Erststimmenergebnis
    treat_mat_caflg <- treat_mat2[,seq(2,10,by=2)]
  } else {
    share_dif_caflg <- share_dif2[,seq(2,10,by=2)]    # H1b: Direktmandat auf Zweitstimmenergebnis
    treat_mat_caflg <- treat_mat2[,seq(1,9,by=2)]
  }
}
```

# Year index - without SPD
```{r}
year2 <- data_now$Jahr
year2 <- rep(year2,ncol(share_dif_caflg))
```

# Party index - without SPD

```{r}
party_caflg <- c("CDU/CSU", "AfD", "FDP", "Linke", "Grüne")
party_caflg <- rep(party_caflg,each=nrow(share_dif_caflg))
```

## Regression - without SPD

```{r}
summary(lm(unlist(share_dif_caflg)~unlist(treat_mat_caflg)  ))
summary(lm(unlist(share_dif_caflg)~unlist(treat_mat_caflg) * as.factor(party_caflg) * as.factor(year2)) )
```



# Relative vote share - without CDU & SPD

```{r}
share_dif3 <- share_dif0[,5:12]  # Which parties? 
```

# Specifying treatments - without CDU & SPD

```{r}
treat_mat3 <- data_treat[,14:23]
treat_mat3[treat_mat3>0] <- 1    # All treatment included

# treat_mat3[treat_mat3==2|treat_mat3==4|treat_mat3==6] <- 0    # only chancellor candidates
# treat_mat3[treat_mat3>0] <- 1    

# treat_mat3[treat_mat3==1|treat_mat3==3|treat_mat3==5] <- 0    # only leading candidates without cc
# treat_mat3[treat_mat3>0] <- 1

# treat_mat3[treat_mat3==1|treat_mat3==2] <- 0    # candidates who have run for direct mandate before
# treat_mat3[treat_mat3>0] <- 1

```

# Data selection - without CDU & SPD

```{r}
which.vote <- 1   # Erststimme or Zweitstimme?
contamination <- 1 # Contamination yes or no?


if (contamination==0){
  if (which.vote==1){
    share_dif_aflg <- share_dif3[,seq(1,7,by=2)]    # H1a: Direktmandat auf Erststimmenergebnis
    treat_mat_aflg <- treat_mat3[,seq(1,7,by=2)]
  } else {
    share_dif_aflg <- share_dif3[,seq(2,8,by=2)]    # H2a: Listenplatz auf Zweitstimmenergebnis
    treat_mat_aflg <- treat_mat3[,seq(2,8,by=2)]
  }                                                 # H2?: Interaktion: Direktkandidatur * Parteiliste auf Zweitstimmen
} else {
  if (which.vote==1){
    share_dif_aflg <- share_dif3[,seq(1,7,by=2)]    # H2b: Liste auf Erststimmenergebnis
    treat_mat_aflg <- treat_mat3[,seq(2,8,by=2)]
  } else {
    share_dif_aflg <- share_dif3[,seq(2,8,by=2)]    # H1b: Direktmandat auf Zweitstimmenergebnis
    treat_mat_aflg <- treat_mat3[,seq(1,7,by=2)]
  }
}
```

# Year index - without CDU & SPD

```{r}
year3 <- data_now$Jahr
year3 <- rep(year3,ncol(share_dif_aflg))
```

# Party index - without CDU & SPD

```{r}
party_aflg <- c("AfD", "FDP", "Linke", "Grüne")
party_aflg <- rep(party_aflg,each=nrow(share_dif_aflg))
```

## Regression - without CDU & SPD

```{r}
summary(lm(unlist(share_dif_aflg)~unlist(treat_mat_aflg)  ))
summary(lm(unlist(share_dif_aflg)~unlist(treat_mat_aflg) * as.factor(party_aflg) * as.factor(year3)) )
```

# Relative vote share - only CDU & SPD

```{r}
share_dif4 <- share_dif0[,1:4]  # Which parties? 
```

# Specifying treatments - only CDU & SPD

```{r}
treat_mat4 <- data_treat[,12:15]
treat_mat4[treat_mat4>0] <- 1    # All treatment included

# treat_mat4[treat_mat4==2|treat_mat4==4|treat_mat4==6] <- 0    # only chancellor candidates
# treat_mat4[treat_mat4>0] <- 1    

# treat_mat4[treat_mat4==1|treat_mat4==3|treat_mat4==5] <- 0    # only leading candidates without cc
# treat_mat4[treat_mat4>0] <- 1

# treat_mat4[treat_mat4==1|treat_mat4==2] <- 0    # candidates who have run for direct mandate before
# treat_mat4[treat_mat4>0] <- 1

```

# Data selection - only CDU & SPD

```{r}
which.vote <- 1   # Erststimme or Zweitstimme?
contamination <- 1 # Contamination yes or no?

if (contamination==0){
  if (which.vote==1){
    share_dif_cs <- share_dif4[,seq(1,3,by=2)]    # H1a: Direktmandat auf Erststimmenergebnis
    treat_mat_cs <- treat_mat4[,seq(1,3,by=2)]
  } else {
    share_dif_cs <- share_dif4[,seq(2,4,by=2)]    # H2a: Listenplatz auf Zweitstimmenergebnis
    treat_mat_cs <- treat_mat4[,seq(2,4,by=2)]
  }                                               # H2?: Interaktion: Direktkandidatur * Parteiliste auf Zweitstimmen
} else {
  if (which.vote==1){
    share_dif_cs <- share_dif4[,seq(1,3,by=2)]    # H2b: Liste auf Erststimmenergebnis
    treat_mat_cs <- treat_mat4[,seq(2,4,by=2)]
  } else {
    share_dif_cs <- share_dif4[,seq(2,4,by=2)]    # H1b: Direktmandat auf Zweitstimmenergebnis
    treat_mat_cs <- treat_mat4[,seq(1,3,by=2)]
  }
}
```

# Year index - only CDU & SPD

```{r}
year4 <- data_now$Jahr
year4 <- rep(year4,ncol(share_dif_cs))
```

# Party index - only CDU & SPD

```{r}
party_cs <- c("CDU/CSU", "SPD")
party_cs <- rep(party_cs,each=nrow(share_dif_cs))
```

## Regression - only CDU & SPD

```{r}
summary(lm(unlist(share_dif_cs)~unlist(treat_mat_cs)  ))
summary(lm(unlist(share_dif_cs)~unlist(treat_mat_cs) * as.factor(party_cs) * as.factor(year4)) )
```


# Relative vote share - without Linke

```{r}
share_dif5 <- share_dif0[,c(1:8,11,12)]  # Which parties? 
```

# Specifying treatments - without Linke

```{r}
treat_mat5 <- data_treat[,c(12:19,22,23)]
treat_mat5[treat_mat5>0] <- 1    # All treatment included

# treat_mat4[treat_mat4==2|treat_mat4==4|treat_mat4==6] <- 0    # only chancellor candidates
# treat_mat4[treat_mat4>0] <- 1    

# treat_mat4[treat_mat4==1|treat_mat4==3|treat_mat4==5] <- 0    # only leading candidates without cc
# treat_mat4[treat_mat4>0] <- 1

# treat_mat4[treat_mat4==1|treat_mat4==2] <- 0    # candidates who have run for direct mandate before
# treat_mat4[treat_mat4>0] <- 1

```

# Data selection - without Linke

```{r}
which.vote <- 1   # Erststimme or Zweitstimme?
contamination <- 1 # Contamination yes or no?


if (contamination==0){
  if (which.vote==1){
    share_dif_csafg <- share_dif5[,seq(1,9,by=2)]    # H1a: Direktmandat auf Erststimmenergebnis
    treat_mat_csafg <- treat_mat5[,seq(1,9,by=2)]
  } else {
    share_dif_csafg <- share_dif5[,seq(2,10,by=2)]   # H2a: Listenplatz auf Zweitstimmenergebnis
    treat_mat_csafg <- treat_mat5[,seq(2,10,by=2)]
  }                                                  # H2?: Interaktion: Direktkandidatur * Parteiliste auf Zweitstimmen
} else {
  if (which.vote==1){
    share_dif_csafg <- share_dif5[,seq(1,9,by=2)]    # H2b: Liste auf Erststimmenergebnis
    treat_mat_csafg <- treat_mat5[,seq(2,10,by=2)]
  } else {
    share_dif_csafg <- share_dif5[,seq(2,10,by=2)]   # H1b: Direktmandat auf Zweitstimmenergebnis
    treat_mat_csafg <- treat_mat5[,seq(1,9,by=2)]
  }
}
```

# Year index - without Linke

```{r}
year5 <- data_now$Jahr
year5 <- rep(year5,ncol(share_dif_csafg))
```

# Party index - without Linke

```{r}
party_csafg <- c("CDU/CSU", "SPD", "AfD", "FDP", "Grüne")
party_csafg <- rep(party_csafg,each=nrow(share_dif_csafg))
```

## Regression - without Linke

```{r}
summary(lm(unlist(share_dif_csafg)~unlist(treat_mat_csafg)  ))
summary(lm(unlist(share_dif_csafg)~unlist(treat_mat_csafg) * as.factor(party_csafg) * as.factor(year5)) )
```






## Tries?

```{r}
Formula.PLM.baseline <- party.voteshare ~ 
```

```{r}
## Anteile Gueltiger Stimmen
 

## Anteile Alle Wahlberechtigte
## 
## 
## aus Treatment Matric verschiedene Matrixen erstellen
## 
## alle SK Kandidaten !, alle anderen null
## 
## Alle nicht Nuller gleich 1
```


```{r}
model1 <- plm()

lm(c(Y ~c(X))
```

