---
title: "Coverage Matrix: Candidates of Interest in Politbarometer & ARD Deutschlandtrend"
author: "Anna Werner"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(knitr)
```

## Overview

This document shows, for each Bundestagswahl year (2002–2025), which **candidates of interest**
(Spitzenkandidaten / Kanzlerkandidaten) are covered in the two survey sources used in this project:

- **Politbarometer (PB):** Skalometer –5 to +5 (GESIS ZA studies). Variable name given where available.
- **ARD Deutschlandtrend (DTrend):** 4-point satisfaction question (sehr/eher/weniger/gar nicht zufrieden).
  Additional column for whether a dedicated **Kanzlerkandidat-/Kanzlerpräferenz-Frage** was asked.

**Legend:**

- ✅ = covered / asked
- ❌ = not covered / not asked
- `KK` = Kanzlerkandidat(in)
- `SK` = Spitzenkandidat(in) (not necessarily the KK)
- `—` = not applicable / no dedicated variable

---

## BTW 2002

**PB:** ZA3849 (East) / ZA3850 (West)  |  **DTrend:** ZA4594

```{r btw2002}
df02 <- data.frame(
  Candidate          = c("Gerhard Schröder", "Edmund Stoiber",
                         "Joschka Fischer", "Guido Westerwelle",
                         "Gabi Zimmer",
                         "Gregor Gysi",
                         "Renate Künast", "Angela Merkel"),
  Party              = c("SPD", "CDU/CSU",
                         "Grüne", "FDP",
                         "PDS",
                         "PDS/Linke",
                         "Grüne", "CDU"),
  Role               = c("KK", "KK",
                         "SK", "SK",
                         "SK",
                         "extra (in PB, not SK)",
                         "extra (in PB, not official SK)", "extra (in PB, opposition)"),
  PB_Variable        = c("v103", "v105",
                         "v94", "v108",
                         "❌ Missing",
                         "v95",
                         "v97 (Künast not listed; Merkel at v97)", "v97"),
  DTrend_Zuf         = c("✅", "✅",
                         "✅", "✅",
                         "✅ (Zimmer directly)",
                         "❌",
                         "❌", "✅"),
  DTrend_KK_Frage    = c("✅ Schröder vs. Stoiber; vs. Merkel", "✅ Schröder vs. Stoiber",
                         "❌", "❌",
                         "❌",
                         "❌",
                         "❌", "✅ Schröder vs. Merkel"),
  stringsAsFactors = FALSE
)
kable(df02, col.names = c("Candidate", "Party", "Role",
                          "PB Variable", "DTrend Zuf.", "DTrend KK-Frage"),
      align = "llllll")
```

**Notes:**

- PDS SK in 2002 was Gabi Zimmer — she is **missing** in the PB Skalometer. Gysi (v95) was included
  in PB but was not the official SK.
- v97 = Merkel (CDU, opposition leader, extra); Künast is also noted as extra in PB codebooks but
  her exact variable is to be confirmed (may not have a separate Skalometer item).
- DTrend 2002: Politikerzufriedenheit (4-stufig) + Direktwahl-KK (Schröder vs. Stoiber; Schröder vs. Merkel).

---

## BTW 2005

**PB:** ZA4258 (East) / ZA4259 (West)  |  **DTrend:** ZA4597

```{r btw2005}
df05 <- data.frame(
  Candidate          = c("Gerhard Schröder", "Angela Merkel",
                         "Joschka Fischer", "Guido Westerwelle",
                         "Oskar Lafontaine", "Gregor Gysi"),
  Party              = c("SPD", "CDU/CSU",
                         "Grüne", "FDP",
                         "Linke", "Linke"),
  Role               = c("KK", "KK",
                         "SK", "SK",
                         "SK", "SK"),
  PB_Variable        = c("v154", "v148",
                         "v143", "v158",
                         "v147", "v144"),
  DTrend_Zuf         = c("✅", "✅",
                         "✅", "✅",
                         "❌", "❌"),
  DTrend_KK_Frage    = c("✅ Schröder vs. Merkel", "✅ Schröder vs. Merkel",
                         "❌", "❌",
                         "❌", "❌"),
  stringsAsFactors = FALSE
)
kable(df05, col.names = c("Candidate", "Party", "Role",
                          "PB Variable", "DTrend Zuf.", "DTrend KK-Frage"),
      align = "llllll")
```

**Notes:**

- All candidates of interest are covered in PB (no missing).
- Lafontaine and Gysi (both Linke SK) are **not** in DTrend Zufriedenheit.
- DTrend 2005: Politikerzufriedenheit + Direktwahl-KK (Schröder vs. Merkel)
  + separate question on Union-KK (Merkel vs. Stoiber).
- PB extra vars: Kanzlerpräferenz.

---

## BTW 2009

**PB:** ZA5431 (East) / ZA5432 (West)  |  **DTrend:** ZA5448

Split sample in PB (East/West) — same variable names for both subsets.

```{r btw2009}
df09 <- data.frame(
  Candidate          = c("Angela Merkel", "Frank-Walter Steinmeier",
                         "Guido Westerwelle",
                         "Jürgen Trittin", "Renate Künast",
                         "Oskar Lafontaine", "Gregor Gysi"),
  Party              = c("CDU/CSU", "SPD",
                         "FDP",
                         "Grüne", "Grüne",
                         "Linke", "Linke"),
  Role               = c("KK", "KK",
                         "SK",
                         "SK", "SK",
                         "SK", "SK"),
  PB_Variable        = c("V109", "V116",
                         "V117",
                         "❌ Missing", "❌ Missing",
                         "V108", "V105"),
  DTrend_Zuf         = c("✅", "✅",
                         "✅",
                         "✅", "❌",
                         "✅", "❌"),
  DTrend_KK_Frage    = c("✅ Merkel vs. Steinmeier", "✅ Merkel vs. Steinmeier",
                         "❌",
                         "❌", "❌",
                         "❌", "❌"),
  stringsAsFactors = FALSE
)
kable(df09, col.names = c("Candidate", "Party", "Role",
                          "PB Variable", "DTrend Zuf.", "DTrend KK-Frage"),
      align = "llllll")
```

**Notes:**

- **Trittin** (Grüne SK) is in DTrend but **missing** in PB.
- **Künast** (Grüne SK) is missing in **both** sources.
- **Gysi** is in PB but **not** in DTrend.
- DTrend 2009: Politikerzufriedenheit + Direktwahl-KK (Merkel vs. Steinmeier, all waves Jan–Sep).

---

## BTW 2013

**PB:** ZA5677  |  **DTrend:** ZA5915

```{r btw2013}
df13 <- data.frame(
  Candidate          = c("Angela Merkel", "Peer Steinbrück",
                         "Rainer Brüderle",
                         "Jürgen Trittin", "Katrin Göring-Eckardt",
                         "Gregor Gysi",
                         "Sahra Wagenknecht",
                         "Dietmar Bartsch", "Jan van Aken", "Klaus Ernst",
                         "Caren Lay", "Nicole Gohlke", "Diana Golze"),
  Party              = c("CDU/CSU", "SPD",
                         "FDP",
                         "Grüne", "Grüne",
                         "Linke",
                         "Linke",
                         "Linke", "Linke", "Linke",
                         "Linke", "Linke", "Linke"),
  Role               = c("KK", "KK",
                         "SK",
                         "SK", "SK",
                         "SK",
                         "SK",
                         "SK", "SK", "SK",
                         "SK", "SK", "SK"),
  PB_Variable        = c("V119", "V124",
                         "V114",
                         "V126", "❌ Missing",
                         "V121",
                         "❌ Missing",
                         "❌ Missing", "❌ Missing", "❌ Missing",
                         "❌ Missing", "❌ Missing", "❌ Missing"),
  DTrend_Zuf         = c("✅", "✅",
                         "✅",
                         "✅", "✅",
                         "✅",
                         "❌",
                         "❌", "❌", "❌",
                         "❌", "❌", "❌"),
  DTrend_KK_Frage    = c("✅ Merkel vs. Steinbrück", "✅ Merkel vs. Steinbrück",
                         "❌",
                         "❌", "❌",
                         "❌",
                         "❌",
                         "❌", "❌", "❌",
                         "❌", "❌", "❌"),
  stringsAsFactors = FALSE
)
kable(df13, col.names = c("Candidate", "Party", "Role",
                          "PB Variable", "DTrend Zuf.", "DTrend KK-Frage"),
      align = "llllll")
```

**Notes:**

- **Göring-Eckardt** (Grüne SK, co-top with Trittin) is in DTrend but **missing** in PB.
- **Wagenknecht** (Linke SK) is missing in **both** sources.
- All other Linke top-list candidates (Bartsch, van Aken, Ernst, Lay, Gohlke, Golze) are missing
  in **both** sources.
- DTrend 2013: Politikerzufriedenheit + Direktwahl-KK (Merkel vs. Steinbrück)
  + Aussagen zu Merkel (Euro-Krise) + Aussagen zu Steinbrück (Kanzlertauglichkeit).

---

## BTW 2017

**PB:** ZA6988  |  **DTrend:** ZA6987

```{r btw2017}
df17 <- data.frame(
  Candidate          = c("Angela Merkel", "Martin Schulz",
                         "Christian Lindner",
                         "Cem Özdemir", "Katrin Göring-Eckardt",
                         "Sahra Wagenknecht", "Dietmar Bartsch",
                         "Alice Weidel", "Alexander Gauland"),
  Party              = c("CDU/CSU", "SPD",
                         "FDP",
                         "Grüne", "Grüne",
                         "Linke", "Linke",
                         "AfD", "AfD"),
  Role               = c("KK", "KK",
                         "SK",
                         "SK", "SK",
                         "SK", "SK",
                         "SK", "SK"),
  PB_Variable        = c("V107", "V111",
                         "V104",
                         "V109", "❌ Missing",
                         "V114", "❌ Missing",
                         "❌ Missing", "❌ Missing"),
  DTrend_Zuf         = c("✅", "✅",
                         "✅",
                         "✅", "❌",
                         "✅", "❌",
                         "❌", "❌"),
  DTrend_KK_Frage    = c("✅ Merkel vs. Schulz", "✅ Merkel vs. Schulz",
                         "❌",
                         "❌", "❌",
                         "❌", "❌",
                         "❌", "❌"),
  stringsAsFactors = FALSE
)
kable(df17, col.names = c("Candidate", "Party", "Role",
                          "PB Variable", "DTrend Zuf.", "DTrend KK-Frage"),
      align = "llllll")
```

**Notes:**

- **Göring-Eckardt** (Grüne co-SK with Özdemir) is missing in **both** sources.
- **Bartsch** (Linke co-SK with Wagenknecht) is missing in **both** sources.
- Both **AfD SK** (Weidel, Gauland) are missing in **both** sources.
- DTrend 2017: Politikerzufriedenheit + Direktwahl-KK (Merkel vs. Schulz) + TV-Duell items.

---

## BTW 2021

**PB:** ZA7856  |  **DTrend:** ZA7863

```{r btw2021}
df21 <- data.frame(
  Candidate          = c("Armin Laschet", "Olaf Scholz",
                         "Annalena Baerbock",
                         "Christian Lindner",
                         "Sahra Wagenknecht",
                         "Janine Wissler", "Dietmar Bartsch",
                         "Alice Weidel", "Tino Chrupalla"),
  Party              = c("CDU/CSU", "SPD",
                         "Grüne",
                         "FDP",
                         "Linke",
                         "Linke", "Linke",
                         "AfD", "AfD"),
  Role               = c("KK", "KK",
                         "KK",
                         "SK",
                         "Fraktionsvors. (not official SK)",
                         "SK", "SK",
                         "SK", "SK"),
  PB_Variable        = c("V110", "V116",
                         "V106",
                         "V112",
                         "V112 ⚠️",
                         "❌ Missing", "❌ Missing",
                         "❌ Missing", "❌ Missing"),
  DTrend_Zuf         = c("✅", "✅",
                         "❌",
                         "✅",
                         "❌",
                         "❌", "❌",
                         "❌", "❌"),
  DTrend_KK_Frage    = c("✅ Laschet vs. Söder (CDU/CSU KK-Frage)", "❌",
                         "❌",
                         "❌",
                         "❌",
                         "❌", "❌",
                         "❌", "❌"),
  stringsAsFactors = FALSE
)
kable(df21, col.names = c("Candidate", "Party", "Role",
                          "PB Variable", "DTrend Zuf.", "DTrend KK-Frage"),
      align = "llllll")
```

**Notes:**

- ⚠️ **V112 appears twice** in the codebook (for both Lindner and Wagenknecht in 2021) — please
  verify in ZA7856; one of these is likely a different variable or a codebook typo.
- **Baerbock** is in PB (V106) but is **missing** in the DTrend January 2021 wave (not yet nominated
  at that point; Grüne nomination came in April 2021).
- Both **Linke SK** (Wissler, Bartsch) and both **AfD SK** (Weidel, Chrupalla) are missing in
  **both** sources.
- PB extra var: partly asked in *Wichtiger Politiker* variable.
- DTrend 2021: Politikerzufriedenheit (Jan 2021 wave only in ZA7863).

---

## BTW 2025

**PB:** Not yet accessible (apply for access at GESIS, expected ZA8827 or similar)  
**DTrend:** ZA9050 (waves Jan–Dec 2024, incl. Nov II & Dez II with KK-specific items)

```{r btw2025}
df25 <- data.frame(
  Candidate          = c("Friedrich Merz", "Olaf Scholz",
                         "Robert Habeck",
                         "Alice Weidel",
                         "Sahra Wagenknecht",
                         "Christian Lindner"),
  Party              = c("CDU/CSU", "SPD",
                         "Grüne",
                         "AfD",
                         "BSW",
                         "FDP"),
  Role               = c("KK", "KK",
                         "KK",
                         "KK",
                         "SK",
                         "SK"),
  PB_Variable        = c("⏳ PB n.v.", "⏳ PB n.v.",
                         "⏳ PB n.v.",
                         "⏳ PB n.v.",
                         "⏳ PB n.v.",
                         "⏳ PB n.v."),
  DTrend_Zuf         = c("✅ (Jan, Nov, Nov II, Dez 2024)", "✅ (Jan–Dez 2024)",
                         "✅ (Jan, Nov–Dez 2024)",
                         "✅ (Nov, Nov II, Dez 2024)",
                         "✅ (Nov, Nov II, Dez 2024)",
                         "✅ (Jan–Nov 2024)"),
  DTrend_KK_Frage    = c("✅ Nov II: KK-Bewertung; Dez II: Kanzler/in-Frage; Kanzlerpräf. Merz vs. Scholz",
                         "✅ Nov II; Dez II: Kanzler/in + Kanzlerpräferenz",
                         "✅ Nov II; Dez II: Kanzler/in-Frage",
                         "✅ Nov II; Dez II: Kanzler/in-Frage",
                         "❌",
                         "❌"),
  stringsAsFactors = FALSE
)
kable(df25, col.names = c("Candidate", "Party", "Role",
                          "PB Variable", "DTrend Zuf.", "DTrend KK-Frage"),
      align = "llllll")
```

**Notes:**

- PB 2025 is not yet available in the repository. Access must be requested from GESIS.
- DTrend 2025: ZA9050 contains 13 waves (Jan–Dez II 2024). From November 2024 onward,
  dedicated Kanzlerkandidaten-Bewertungsfragen are included for all 4 main KK.
- DTrend scale is 4-point ordinal — **not** directly comparable to PB Skalometer (–5 to +5).

---

## Summary: All Candidates of Interest

```{r summary}
summary_df <- data.frame(
  Year       = c(2002, 2002, 2002, 2002, 2002,
                 2005, 2005, 2005, 2005, 2005, 2005,
                 2009, 2009, 2009, 2009, 2009, 2009, 2009,
                 2013, 2013, 2013, 2013, 2013, 2013, 2013,
                 2017, 2017, 2017, 2017, 2017, 2017, 2017, 2017, 2017,
                 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021,
                 2025, 2025, 2025, 2025, 2025, 2025),
  Candidate  = c("Schröder", "Stoiber", "Fischer", "Westerwelle", "Zimmer",
                 "Schröder", "Merkel", "Fischer", "Westerwelle", "Lafontaine", "Gysi",
                 "Merkel", "Steinmeier", "Westerwelle", "Trittin", "Künast", "Lafontaine", "Gysi",
                 "Merkel", "Steinbrück", "Brüderle", "Trittin", "Göring-Eckardt", "Gysi", "Wagenknecht",
                 "Merkel", "Schulz", "Lindner", "Özdemir", "Göring-Eckardt", "Wagenknecht", "Bartsch", "Weidel", "Gauland",
                 "Laschet", "Scholz", "Baerbock", "Lindner", "Wagenknecht", "Wissler", "Bartsch", "Weidel", "Chrupalla",
                 "Merz", "Scholz", "Habeck", "Weidel", "Wagenknecht", "Lindner"),
  Party      = c("SPD", "CDU/CSU", "Grüne", "FDP", "PDS",
                 "SPD", "CDU/CSU", "Grüne", "FDP", "Linke", "Linke",
                 "CDU/CSU", "SPD", "FDP", "Grüne", "Grüne", "Linke", "Linke",
                 "CDU/CSU", "SPD", "FDP", "Grüne", "Grüne", "Linke", "Linke",
                 "CDU/CSU", "SPD", "FDP", "Grüne", "Grüne", "Linke", "Linke", "AfD", "AfD",
                 "CDU/CSU", "SPD", "Grüne", "FDP", "Linke", "Linke", "Linke", "AfD", "AfD",
                 "CDU/CSU", "SPD", "Grüne", "AfD", "BSW", "FDP"),
  Role       = c("KK", "KK", "SK", "SK", "SK",
                 "KK", "KK", "SK", "SK", "SK", "SK",
                 "KK", "KK", "SK", "SK", "SK", "SK", "SK",
                 "KK", "KK", "SK", "SK", "SK", "SK", "SK",
                 "KK", "KK", "SK", "SK", "SK", "SK", "SK", "SK", "SK",
                 "KK", "KK", "KK", "SK", "SK (Frakt.)", "SK", "SK", "SK", "SK",
                 "KK", "KK", "KK", "KK", "SK", "SK"),
  PB_Skalo   = c("✅ v103", "✅ v105", "✅ v94", "✅ v108", "❌",
                 "✅ v154", "✅ v148", "✅ v143", "✅ v158", "✅ v147", "✅ v144",
                 "✅ V109", "✅ V116", "✅ V117", "❌", "❌", "✅ V108", "✅ V105",
                 "✅ V119", "✅ V124", "✅ V114", "✅ V126", "❌", "✅ V121", "❌",
                 "✅ V107", "✅ V111", "✅ V104", "✅ V109", "❌", "✅ V114", "❌", "❌", "❌",
                 "✅ V110", "✅ V116", "✅ V106", "✅ V112", "✅ V112 ⚠️", "❌", "❌", "❌", "❌",
                 "⏳", "⏳", "⏳", "⏳", "⏳", "⏳"),
  DTrend_Zuf = c("✅", "✅", "✅", "✅", "✅",
                 "✅", "✅", "✅", "✅", "❌", "❌",
                 "✅", "✅", "✅", "✅", "❌", "✅", "❌",
                 "✅", "✅", "✅", "✅", "✅", "✅", "❌",
                 "✅", "✅", "✅", "✅", "❌", "✅", "❌", "❌", "❌",
                 "✅", "✅", "❌", "✅", "❌", "❌", "❌", "❌", "❌",
                 "✅", "✅", "✅", "✅", "✅", "✅"),
  DTrend_KK  = c("✅", "✅", "❌", "❌", "❌",
                 "✅", "✅", "❌", "❌", "❌", "❌",
                 "✅", "✅", "❌", "❌", "❌", "❌", "❌",
                 "✅", "✅", "❌", "❌", "❌", "❌", "❌",
                 "✅", "✅", "❌", "❌", "❌", "❌", "❌", "❌", "❌",
                 "✅", "❌", "❌", "❌", "❌", "❌", "❌", "❌", "❌",
                 "✅", "✅", "✅", "✅", "❌", "❌"),
  stringsAsFactors = FALSE
)

kable(summary_df, col.names = c("Year", "Candidate", "Party", "Role",
                                "PB Skalometer", "DTrend Zuf.", "DTrend KK-Frage"),
      align = "clllccc")
```

**Systematic gaps (candidate missing in both PB and DTrend):**

| Year | Candidate | Party | Role |
|---|---|---|---|
| 2002 | Zimmer | PDS | SK |
| 2009 | Künast | Grüne | SK |
| 2013 | Göring-Eckardt | Grüne | SK (DTrend only, PB missing) |
| 2013 | Wagenknecht | Linke | SK |
| 2013 | Bartsch, van Aken, Ernst, Lay, Gohlke, Golze | Linke | SK |
| 2017 | Göring-Eckardt | Grüne | SK |
| 2017 | Bartsch | Linke | SK |
| 2017 | Weidel, Gauland | AfD | SK |
| 2021 | Wissler, Bartsch | Linke | SK |
| 2021 | Weidel, Chrupalla | AfD | SK |

---

## Methodological Notes

| Feature | Politbarometer | ARD Deutschlandtrend |
|---|---|---|
| **Scale** | Skalometer –5 to +5 (11-point) | 4-point ordinal (sehr/eher/weniger/gar nicht zufrieden) |
| **Comparability** | Directly comparable across all PB years | Not directly comparable with PB without transformation |
| **Person coverage** | Mainly governing parties + KK | Broader (also ministers, AfD from 2017) |
| **Survey period** | Monthly, full year | Monthly, full year (ZA9050: 13 waves) |
| **Sample size** | ~1,000/wave | ~1,000 (2002–2021); ~1,300 (2024, CATI+Online) |
| **Kanzlerpräferenz** | ✅ (own variable) | ✅ (as separate question) |
| **KK evaluation (gut/nicht gut)** | ❌ | ✅ from Nov 2024 (BTW 2025) |

---

*Sources: ZA3849/3850 (2002), ZA4258/4259 (2005), ZA5431/5432 (2009), ZA5677 (2013),
ZA6988 (2017), ZA7856 (2021) — Politbarometer codebooks (GESIS).  
ZA4594 (2002), ZA4597 (2005), ZA5448 (2009), ZA5915 (2013), ZA6987 (2017),
ZA7863 (2021), ZA9050 (2025) — ARD Deutschlandtrend codebooks (GESIS).*
