---
title: "Residual Analysis: Politician Approval and Treatment Effects"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, message=FALSE, warning=FALSE}
library(haven)
library(readxl)
```

# Overview

The FD model estimates a single pooled treatment effect across all politicians and years. This analysis asks whether the deviation from that average relates to how popular a politician was at the time of the election.

The outcome is sign(treat) x u-hat: positive means the constituency moved more than the model average in the expected direction, negative means less. For PR candidates this is averaged across constituencies. The approval measure is the mean Skalometer rating (1 to 11) from the Politbarometer pre-election wave.

Coverage is 2002 to 2021. There is no 2025 data. AfD politicians are not covered by the Politbarometer. Some Die Linke candidates in 2013 are also missing.

On the correlation: Spearman's rho is used rather than Pearson. With N around 15 to 25 politician-year observations, we cannot assume bivariate normality, and Spearman makes no such assumption. It ranks both variables and correlates the ranks, so it is also robust to the outliers that are likely with this sample size.

# 1. Data loading

```{r load_data, message=FALSE}
load("Data/analysis_dat.RData")

result.files <- sort(list.files(
  "BTW_2025/Outputs/First_Differencing/",
  pattern   = "^party\\.selec1_denom1_\\d{4}-\\d{2}-\\d{2}\\.RData$",
  full.names = TRUE
))
if (length(result.files) == 0) stop("No result files found. Run Analyse_FD_25.Rmd first.")
cat("Loading:", basename(result.files[length(result.files)]), "\n")
load(result.files[length(result.files)])

treat.raw <- read_xlsx("Data/Treatment_WK.xlsx", sheet = 2)
treat.raw  <- treat.raw[!is.na(treat.raw$Jahr), ]
treat.raw  <- treat.raw[!is.na(treat.raw$Partei2), ]
treat.raw$WK <- suppressWarnings(as.numeric(treat.raw$WK))
```

# 2. Politbarometer approval ratings

The Skalometer runs from 1 (very negative) to 11 (very positive). We use the September wave in each year, which is closest to the election. No monthly pooling is done: the mean is computed from the election month only. Where a politician does not appear in September (the questionnaire rotates which politicians are asked about), we fall back to the single latest available pre-election month.

For years where the Politbarometer was fielded as separate east and west samples (2002, 2005, 2009), we combine both files before computing the mean so that the national average is weighted by actual sample size rather than giving equal weight to each half.

For treat_fd equal to -1 (a politician who left), we use the approval rating from the previous election year, since that is the popularity the party lost.

```{r pb_config}
prev.year.map <- c(`2021`=2017, `2017`=2013, `2013`=2009,
                   `2009`=2005, `2005`=2002, `2002`=NA_real_)

pb.config <- list(
  `2021` = list(
    type = "combined",
    file = "Data/politbarometer/ZA7856_v1-1-0_2021.sav",
    month.var = "V4", election.month = 9,
    vars = c(
      "Baerbock, Annalena" = "V106",
      "Laschet, Armin"     = "V110",
      "Lindner, Chrisitan" = "V112",
      "Scholz, Olaf"       = "V116",
      "Wagenknecht, Sarah" = "V122"
    )
  ),
  `2017` = list(
    type = "combined",
    file = "Data/politbarometer/ZA6988_v1-2-0_2017.sav",
    month.var = "V4", election.month = 9,
    vars = c(
      "Lindner, Christian"  = "V104",
      "Merkel, Angela"      = "V107",
      "Özdemir, Cem"        = "V109",
      "Schulz, Martin"      = "V111",
      "Wagenknecht, Sarah"  = "V114"
    )
  ),
  `2013` = list(
    type = "combined",
    file = "Data/politbarometer/ZA5677_v1-1-0_2013.sav",
    month.var = "V4", election.month = 9,
    vars = c(
      "Brüderle, Rainer" = "V114",
      "Merkel, Angela"   = "V119",
      "Gysi, Gregor"     = "V121",
      "Steinbrück, Peer" = "V124",
      "Trittin, Jürgen"  = "V126"
    )
  ),
  `2009` = list(
    type = "split",
    file.west = "Data/politbarometer/ZA5432_v1-2-0_2009_west.sav",
    file.east = "Data/politbarometer/ZA5431_v1-2-0_2009_east.sav",
    month.var = "V4", election.month = 9,
    vars = c(
      "Gysi, Gregor"       = "V105",
      "Lafontaine, Oskar"  = "V108",
      "Merkel, CDU"        = "V109",
      "Steinmeiner, SPD"   = "V116",
      "Westerwelle, Guido" = "V117"
    )
  ),
  `2005` = list(
    type = "split",
    file.west = "Data/politbarometer/ZA4259_v1-1-0_2005_west.sav",
    file.east = "Data/politbarometer/ZA4258_v1-1-0_2005_east.sav",
    month.var = "v3", election.month = 9,
    vars = c(
      "Fischer, Joschka"   = "v143",
      "Gysi, Gregor"       = "v144",
      "Lafontaine, Oskar"  = "v147",
      "Merkel, CDU"        = "v148",
      "Schröder, SPD"     = "v154",
      "Westerwelle, Guido" = "v158"
    )
  ),
  `2002` = list(
    type = "split",
    file.west = "Data/politbarometer/ZA3849_2002_west.sav",
    file.east = "Data/politbarometer/ZA3850_2002_east.sav",
    month.var = "v3", election.month = 9,
    vars = c(
      "Fischer, Joschka"   = "v94",
      "Schröder, SPD"     = "v103",
      "Stoiber, CDU"       = "v105",
      "Westerwelle, Guido" = "v108"
    )
  )
)

extract.mean.rating <- function(dat, vars, month.var, election.month) {
  sapply(vars, function(v) {
    vals <- as.numeric(dat[[v]][dat[[month.var]] == election.month])
    if (any(!is.na(vals))) return(mean(vals, na.rm = TRUE))
    pre <- dat[dat[[month.var]] < election.month & !is.na(dat[[v]]), ]
    if (nrow(pre) == 0) return(NA_real_)
    m <- max(pre[[month.var]], na.rm = TRUE)
    cat("  Note:", v, "month", election.month, "empty; using month", m, "\n")
    mean(as.numeric(pre[[v]][pre[[month.var]] == m]), na.rm = TRUE)
  })
}

build.approval.tab <- function(cfg) {
  rows <- list()
  for (yr in names(cfg)) {
    c. <- cfg[[yr]]
    dat <- if (c.$type == "combined") {
      zap_labels(read_sav(c.$file))
    } else {
      rbind(zap_labels(read_sav(c.$file.west)),
            zap_labels(read_sav(c.$file.east)))
    }
    r <- extract.mean.rating(dat, c.$vars, c.$month.var, c.$election.month)
    for (nm in names(c.$vars))
      rows[[length(rows)+1]] <- data.frame(
        Personen = nm, Jahr = as.numeric(yr), approval = r[nm],
        stringsAsFactors = FALSE)
  }
  do.call(rbind, rows)
}

approval.tab <- build.approval.tab(pb.config)
```

# 3. Directed residuals

```{r directed_residuals}
party.sub <- c(SPD="SPD", CDU="CDU", GRU="Grünen",
               FDP="FDP", AFD="AfD", LIN="Linke", BSW="BSW")

lookup.politician <- function(wkr.nr, party.code, year.t, treat.fd) {
  yr <- if (treat.fd == 1) year.t else prev.year.map[as.character(year.t)]
  if (is.na(yr)) return(NA_character_)
  cands <- treat.raw[
    treat.raw$Jahr == yr &
    grepl(party.sub[party.code], treat.raw$Partei2, ignore.case = TRUE) &
    (as.numeric(treat.raw$T_DM) != 0 | as.numeric(treat.raw$T_LP) != 0), ]
  hit <- cands[!is.na(cands$WK) & cands$WK == wkr.nr, ]
  if (nrow(hit)   > 0) return(hit$Personen[1])
  if (nrow(cands) > 0) return(cands$Personen[1])
  NA_character_
}

model.sets <- list(
  list(res = all.results.smd, tier = "FPTP"),
  list(res = all.results.pr,  tier = "PR")
)

rows.direct <- list()
for (ms in model.sets) {
  lm.dat          <- ms$res[[3]]
  lm.dat$resid    <- residuals(ms$res[[1]])
  lm.dat$year.num <- as.numeric(as.character(lm.dat$year))
  lm.dat$party.ch <- as.character(lm.dat$party)
  lm.dat$tier     <- ms$tier

  dir.obs <- lm.dat[lm.dat$treat != 0, ]
  if (nrow(dir.obs) == 0) next

  dir.obs$directed.effect <- sign(dir.obs$treat) * dir.obs$resid
  dir.obs$Personen <- mapply(lookup.politician,
    wkr.nr     = dir.obs$wkr.nr,
    party.code = dir.obs$party.ch,
    year.t     = dir.obs$year.num,
    treat.fd   = dir.obs$treat)
  dir.obs$approval.year <- ifelse(
    dir.obs$treat == 1, dir.obs$year.num,
    prev.year.map[as.character(dir.obs$year.num)])
  rows.direct[[length(rows.direct)+1]] <- dir.obs
}
dat.direct <- do.call(rbind, rows.direct)
```

# 4. Merge and aggregate

```{r merge_aggregate}
m <- merge(dat.direct, approval.tab,
           by.x = c("Personen", "approval.year"),
           by.y = c("Personen", "Jahr"),
           all.x = TRUE)

make.agg <- function(dat) {
  a    <- aggregate(directed.effect ~ Personen + approval.year + tier, data = dat, FUN = mean)
  appr <- aggregate(approval ~ Personen + approval.year + tier, data = dat,
                    FUN = function(x) mean(x, na.rm = TRUE), na.action = na.pass)
  n    <- aggregate(directed.effect ~ Personen + approval.year + tier, data = dat, FUN = length)
  names(n)[4] <- "n.wkr"
  a <- merge(a, appr, by = c("Personen", "approval.year", "tier"), all.x = TRUE)
  a <- merge(a,    n, by = c("Personen", "approval.year", "tier"))
  a$short.label <- paste0(sub(",.*", "", a$Personen), " '",
                           substr(a$approval.year, 3, 4))
  a
}

agg.direct <- make.agg(m[!is.na(m$approval) & !is.na(m$Personen), ])
agg.all    <- make.agg(m[!is.na(m$Personen), ])
```

# 5. Results

## 5.1 Directed residuals per politician

```{r residual_table}
for (t in c("FPTP", "PR")) {
  sub <- agg.all[agg.all$tier == t, ]
  sub <- sub[order(sub$directed.effect, decreasing = TRUE), ]
  cat("\n", t, "\n")
  print(data.frame(
    politician      = sub$short.label,
    directed.effect = round(sub$directed.effect, 4),
    approval        = round(sub$approval, 2),
    n.wkr           = sub$n.wkr
  ), row.names = FALSE)
}
```

```{r residual_dotplot, fig.width=11, fig.height=5}
par(mfrow = c(1, 2), mar = c(9, 4, 3, 1))

for (t in c("FPTP", "PR")) {
  sub      <- agg.all[agg.all$tier == t, ]
  sub      <- sub[order(sub$directed.effect), ]
  has.appr <- !is.na(sub$approval)
  pt.col   <- ifelse(!has.appr, "grey60",
                     ifelse(sub$directed.effect >= 0, "steelblue", "tomato3"))
  pt.pch   <- ifelse(has.appr, 16, 1)

  plot(seq_along(sub$directed.effect), sub$directed.effect,
       type = "n", xaxt = "n", xlab = "",
       ylab = "sign(treat) x u-hat",
       main = t,
       xlim = c(0.5, nrow(sub) + 0.5))
  segments(seq_along(sub$directed.effect), 0, seq_along(sub$directed.effect),
           sub$directed.effect, col = pt.col, lwd = 1.2)
  points(seq_along(sub$directed.effect), sub$directed.effect,
         pch = pt.pch, col = pt.col, cex = 1.1)
  abline(h = 0, lty = 2, col = "grey40")
  axis(1, at = seq_along(sub$directed.effect), labels = sub$short.label,
       las = 2, cex.axis = 0.65)
}

par(mfrow = c(1, 1))
legend("bottomright", inset = 0.01,
       legend = c("positive (with approval)", "negative (with approval)", "no approval data"),
       pch = c(16, 16, 1), col = c("steelblue", "tomato3", "grey60"),
       cex = 0.75, bty = "n")
```

## 5.2 Approval and directed residual

```{r scatter, fig.width=10, fig.height=5}
par(mfrow = c(1, 2), mar = c(4, 4, 3, 1))

for (t in c("FPTP", "PR")) {
  sub <- agg.direct[agg.direct$tier == t, ]

  plot(sub$approval, sub$directed.effect,
       pch = 16, cex = 0.9, col = "grey30",
       xlab = "Skalometer (1 to 11)",
       ylab = "sign(treat) x u-hat",
       main = t,
       xlim = c(2, 11))
  abline(h = 0, lty = 2, col = "grey60")
  text(sub$approval, sub$directed.effect,
       labels = sub$short.label, cex = 0.6, pos = 3, col = "grey40")

  if (nrow(sub) >= 3) {
    rho <- cor(sub$approval, sub$directed.effect,
               method = "spearman", use = "complete.obs")
    legend("topleft", bty = "n", cex = 0.85,
           legend = paste0("Spearman rho = ", round(rho, 2),
                           "  (N = ", nrow(sub), ")"))
  }
}

par(mfrow = c(1, 1))
```

N at the politician-year level is small (around 15 to 25 for FPTP, more for PR). The results should be read as exploratory. AfD politicians are not in the Politbarometer. The 2025 election is not covered.
