---
title: "Visualization_FD_Results"
author: "Susumu Shikano"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



### Figures

```{r}
library(readxl)

results <- read_xlsx("Data/ModelSummaries.xlsx",sheet=3)

```
## Effect size with confidence intervals

### Figures in the BA-thesis 

```{r,warning=F,echo=FALSE}
# Define color palette
pacman::p_load(RColorBrewer)

party.colors <- #brewer.pal(9, "Set2")
                #c("#1E366F","#6D1E1E", "#6F6D1E", "#2B2B2B", "#B43104","#B5A200",                          "#993366","#3E6F2D", "#005F6B")
                c("#1D406E","#6E1D1D", "#6F6D1D", "#2B2B2B", "#B43004",
                  "#B59F00", "#9F3A62","#3E6E2C", "#005F6A")

# Add LaTeX font
library(extrafont)
loadfonts()
library(fontcm)


for (i.vote.d in 1:2){# depvar is FPTP or PR
  for (i.vote.i in 1:2){# indepvar is FPTP or PR candidacy

    # Filter data by voting scenario
    data.tv <- as.data.frame(results[results$vote_d==i.vote.d & results$vote_i==i.vote.i & !is.na(results$vote_i),])
    rownames(data.tv) <- data.tv$Party
    
    # Calculate estimate, standard error, and confidence intervals
    p.estimate <- as.numeric(data.tv$Estimate)
    se <- as.numeric(data.tv$SE)                      # not the robust SE!!!
    lower.b <- p.estimate - 1.96* se
    upper.b <- p.estimate + 1.96* se
    
    # Set y-axis range based on confidence intervals
    y.range <- range(c(lower.b,upper.b))
    
    # Combine data into one data frame for visualization
    data.2 <- as.data.frame(cbind(p.estimate, se, lower.b, upper.b))
    rownames(data.2) <- data.tv$Party
    data.vis <- cbind(data.tv, data.2)
    
    # Set title based on voting scenario
    if (i.vote.d==1 & i.vote.i==1) title <- "Direktkandidatur auf Erststimmenergebnis"
    if (i.vote.d==1 & i.vote.i==2) title <- "Listenpr채senz auf Erststimmenergebnis"
    if (i.vote.d==2 & i.vote.i==1) title <- "Direktkandidatur auf Zweitstimmenergebnis"
    if (i.vote.d==2 & i.vote.i==2) title <- "Listenpr채senz auf Zweitstimmenergebnis"
    
    # Create PDF file for the current voting scenario
    pdf(file=paste0("Figures/",title,".pdf"),
        width=12,
        height=9,
        pointsize = 13,
        family = "CM Roman")
    
    # Plot with colors and labels
    plot(p.estimate, axes=F,
       xlab="Spitzenkandidaten der Parteien", 
       ylab="Unterschied im Stimmenanteil",
       ylim=y.range,
    #main=title,
    col=party.colors, lwd = 1.5)

    # Add vertical lines at each data point
    for (i.party in 1:length(p.estimate)){
    lines(c(i.party,i.party),
          c(lower.b[i.party],upper.b[i.party]),
            col=party.colors[i.party], lwd=1.5)
    points(i.party, p.estimate[i.party], col = "black", pch = 21)
    }

    # Depending on voting scenario, manually add horizontal help lines
    # abline(h = 0, lty = 1)          # for all (optional)
    # abline(h = 0.002, lty = 2)      # "LP auf 1"
    # abline(h = 0.0025, lty = 2)     # "DK auf 2"
    # abline(h = 0.004, lty = 2)      # "LP auf 1"
    # abline(h = 0.005, lty = 2)      # "DK auf 2" 
                                      # "LP auf 2"
    # abline(h = 0.006, lty = 2)      # "LP auf 1"
    # abline(h = 0.0075, lty = 2)     # "DK auf 2" 
                                      # "LP auf 2"
     #abline(h = 0.01, lty = 2)       # "DK auf 1"
                                      # "LP auf 2"
    # abline(h = 0.015, lty = 2)      # "DK auf 1"
    # abline(h = 0.0175, lty = 2)     # "LP auf 2"
     abline(h = 0.02, lty = 2)       # "DK auf 1" 
                                      # "LP auf 2"
    # abline(h = 0.012, lty = 2)      # "LP auf 1" 
    
    # Add x-axis and y-axis labels
    axis(1, at=c(1:9), data.tv$Party, cex.axis=0.7)
    axis(2)
    
    # Add a frame
    box()

    dev.off() # close the current PDF file
    #embed_fonts(paste0(title,".pdf"), outfile=paste0(title,".pdf"))
    
    
    if (i.vote.d==1 & i.vote.i==1) data.vis.H1a <- data.vis
    if (i.vote.d==1 & i.vote.i==2) data.vis.H1b <- data.vis
    if (i.vote.d==2 & i.vote.i==2) data.vis.H2a <- data.vis
    if (i.vote.d==2 & i.vote.i==1) data.vis.H2b <- data.vis
 }
}
```


### Figures in the MPSA paper

Results of FD analysis are stored in a list with length of 5:
  1: lm.out
  2: lm.out.robust
  3: lm.dat
  4: votes.in.array
  5: r.se

We extract the following relevant information and create a vector with length of 5, respectively:

  1: FPTP with all parties
  2: PR with all parties
  3: FPTP with major parties
  4: PR with major parties
  5: FPTP with minor parties
  6: PR with minor parties


```{r}
all.results <- NULL

for (i.switch in 1:3){
  load(paste0("D:/Nextcloud/R-Programme_Sync/Anna_BA/Data/Estimation_Results_PSelect",
              i.switch,
              "_denom1.RData"))  
  for (i.tier in 1:2){
    if (i.tier==1) {
      temp.results <- all.results.smd
      } else {
      temp.results <- all.results.pr
      }
      rel.information <- temp.results[[2]][2:3,1:2]
      treat.size <- apply(temp.results[[3]][,2:3]!=0,2,sum)
      treat.size <- rbind(treat.size,treat.size)
      rel.information <- cbind(i.switch,i.tier,rel.information,treat.size)
      all.results <- rbind(all.results,rel.information)
  }
}
all.results[rownames(all.results)=="cross.treat","treat"] <- 
             all.results[rownames(all.results)=="cross.treat","cross.treat"]

all.results <- as.data.frame(all.results)
```




```{r,warning=F}

# Sort data 
    data.tv <- all.results
    data.tv <- rbind(
        data.tv[grep("cross",rownames(data.tv),invert = T),],
        data.tv[grep("cross",rownames(data.tv),invert = F),]
    )
    data.tv <- data.tv[order(data.tv$i.tier),]
    
    # Calculate estimate, standard error, and confidence intervals
    data.tv$lower.b <- data.tv$Estimate - 1.96* data.tv$`Std. Error`
    data.tv$upper.b <- data.tv$Estimate + 1.96* data.tv$`Std. Error`
    
    # Set y-axis range based on confidence intervals
    y.range <- range(c(data.tv$lower.b,data.tv$upper.b))
    y.range[2] <- y.range[2]+0.02
    y.range[1] <- y.range[1]-0.01
    
    data.tv <- rbind(data.tv[1:3,],NA,data.tv[4:6,],NA,
                     data.tv[7:9,],NA,data.tv[10:12,])
        

    title <- "All_Estimated_Effects"
    # Create PDF file for the current voting scenario
    pdf(file=paste0("Figures/",title,".pdf"),
        width=8,
        height=5.5,
        pointsize = 13,
        family = "CM Roman")

    
    # Plot with colors and labels
    plot(data.tv$Estimate, axes=F,
       xlab="", 
       ylab="Effect on vote share",
       ylim=y.range,
       xlim=c(0,nrow(data.tv)+1),
    #main=title,
    pch=c(1:3)[data.tv$i.switch],
    lwd = 1.5)

    data.tv$i.switch[is.na(data.tv$i.switch)] <- 0

    # Add vertical lines at each data point
    for (i.party in 1:nrow(data.tv)){
    lines(c(i.party,i.party),
          c(data.tv$lower.b[i.party],data.tv$upper.b[i.party]),
                lty=1, lwd=2.5)
    #points(i.party, p.estimate[i.party], col = "black", pch = 21)
    }

    # Add x-axis and y-axis labels
    axis(1, label=F,at=c(1:3,5:7,9:11,13:15))
    axis(2)
    abline(h=0,lty=2)

    data.tv$i.switch[data.tv$i.switch==0] <- NA
text(x = 1:nrow(data.tv),
     ## Move labels to just below bottom of chart.
     y = par("usr")[3] - 0.005,
     ## Use names from the data list.
     labels = c("All","Major","Minor")[data.tv$i.switch],
     ## Change the clipping region.
     xpd = NA,
     ## Rotate the labels by 35 degrees.
     srt = 35,
     ## Adjust the labels to 80% right-justified.
     adj = 0.8,
     ## Increase label size.
     cex = 1.1)
  text(4,0.06,"Effects on FPTP vote share",pos=3)
  text(12,0.06,"Effects on PR vote share",pos=3)
  text(2,0.06,"FPTP candidacy",pos=1)
  text(6,0.06,"On the party list",pos=1)
  text(10,0.06,"On the party list",pos=1)
  text(14,0.06,"FPTP candidacy",pos=1)

    #data.tv$treat[is.na(data.tv$treat)] <- ""
  text(c(1:15)[!is.na(data.tv$treat)],rep(-0.02,15)[!is.na(data.tv$treat)],
       paste0("(",data.tv$treat,")")[!is.na(data.tv$treat)],
       cex=0.8)
    dev.off() # close the current PDF file
    #embed_fonts(paste0(title,".pdf"), outfile=paste0(title,".pdf"))
    
    
```

```{r}
# Create data
dat.Ha <- as.data.frame(rbind(data.vis.H1a[1,23], data.vis.H2a[1,23], data.vis.H1a[2,23], data.vis.H2a[2,23],data.vis.H1a[3,23], data.vis.H2a[3,23]))
names(dat.Ha) <- "p.estimate"
p.estimate <- dat.Ha$p.estimate
model <- dat.Ha$model <- c(1,1,2,2,3,3)
party <- dat.Ha$party <- c("Alle Parteien", "Alle Parteien", "CDU/CSU und SPD", "CDU/CSU und SPD", "Kleinere Parteien", "Kleinere Parteien")
hyp <- dat.Ha$hyp <- c("1a","2a","1a","2a","1a","2a")
se <- dat.Ha$se <- c(data.vis.H1a[1,24], data.vis.H2a[1,24], data.vis.H1a[2,24], data.vis.H2a[2,24],data.vis.H1a[3,24], data.vis.H2a[3,24])
lower.b <- dat.Ha$lower.b <- c(data.vis.H1a[1,25], data.vis.H2a[1,25], data.vis.H1a[2,25], data.vis.H2a[2,25],data.vis.H1a[3,25], data.vis.H2a[3,25])
upper.b <- dat.Ha$upper.b <- c(data.vis.H1a[1,26], data.vis.H2a[1,26], data.vis.H1a[2,26], data.vis.H2a[2,26],data.vis.H1a[3,26], data.vis.H2a[3,26])

y.range <- range(c(lower.b,upper.b))


party.colors <- c("#1D406E","#1D406E","#6E1D1D","#6E1D1D","#6F6D1D","#6F6D1D")

# pdf
pdf(file= "H3a_graphix.pdf",
        width=12,
        height=9,
        pointsize = 13,
        family = "CM Roman")

# Plot with colors and labels
plot(p.estimate, axes=F,
    xlab="Spitzenkandidaten der Parteien", ylab="Unterschied im Stimmenanteil",
    ylim=y.range,
    #main="Der Spitzenkandidateneffekt: Direktandidatur (je links) vs. Listenpr채senz (je rechts)",
    col=party.colors, lwd = 1.5)

# Add vertical lines at each data point
    for (i.party in 1:6){
    lines(c(i.party,i.party),
          c(lower.b[i.party],upper.b[i.party]),
            col=party.colors[i.party], lwd=1.5)
    points(i.party, p.estimate[i.party], col = "black", pch = 21)
}

# Depending on voting scenario, manually add horizontal help lines
  #abline(h = 0, lty = 1)
  #abline(h = 0.0038, lty = 2)
  #abline(h = 0.0070, lty = 2)
  #abline(h = 0.0139, lty = 2)
  #abline(h = 0.0145, lty = 2)
  #abline(h = 0.0179, lty = 2)
  #abline(h = 0.0186, lty = 2)
     
# Add x-axis and y-axis labels
  axis(1, at=c(1:6), dat.Ha$party, cex.axis=0.7)
  axis(2)
    
# Add a frame
  box()

  dev.off() # close the current PDF file
```



```{r}
dat.Hb <- as.data.frame(rbind(data.vis.H1b[1,23], data.vis.H2b[1,23], data.vis.H1b[2,23], data.vis.H2b[2,23], data.vis.H1b[3,23], data.vis.H2b[3,23]))
names(dat.Hb) <- "p.estimate"
p.estimate <- dat.Hb$p.estimate
model <- dat.Hb$model <- c(1,1,2,2,3,3)
party <- dat.Hb$party <- c("Alle Parteien", "Alle Parteien", "CDU/CSU und SPD", "CDU/CSU und SPD", "Kleinere Parteien", "Kleinere Parteien")
hyp <- dat.Hb$hyp <- c("1b","2b","1b","2b","1b","2b")
se <- dat.Hb$se <- c(data.vis.H1b[1,24], data.vis.H2b[1,24], data.vis.H1b[2,24], data.vis.H2b[2,24], data.vis.H1b[3,24], data.vis.H2b[3,24])
lower.b <- dat.Hb$lower.b <- c(data.vis.H1b[1,25], data.vis.H2b[1,25], data.vis.H1b[2,25], data.vis.H2b[2,25], data.vis.H1b[3,25], data.vis.H2b[3,25])
upper.b <- dat.Hb$upper.b <- c(data.vis.H1b[1,26], data.vis.H2b[1,26], data.vis.H1b[2,26], data.vis.H2b[2,26], data.vis.H1b[3,26], data.vis.H2b[3,26])


y.range <- range(c(lower.b,upper.b))

party.colors <- c("#1D406E","#1D406E","#6E1D1D","#6E1D1D","#6F6D1D","#6F6D1D")

# pdf
pdf(file= "H3b_graphix.pdf",
        width=12,
        height=9,
        pointsize = 13,
        family = "CM Roman")

# Plot with colors and labels
plot(p.estimate, axes=F,
    xlab="Spitzenkandidaten der Parteien", ylab="Unterschied im Stimmenanteil",
    ylim=y.range,
    #main="Kontaminationseffekte: Direktandidatur (je links) vs. Listenpr채senz (je rechts)",
    col=party.colors, lwd = 1.5)

# Add vertical lines at each data point
    for (i.party in 1:6){
    lines(c(i.party,i.party),
          c(lower.b[i.party],upper.b[i.party]),
            col=party.colors[i.party], lwd=1.5)
    points(i.party, p.estimate[i.party], col = "black", pch = 21)
}

    # Depending on voting scenario, manually add horizontal help lines
     abline(h = 0, lty = 1)
     abline(h = 0.0042, lty = 2)
     abline(h = 0.0060, lty = 2)
     abline(h = 0.0114, lty = 2)
     abline(h = 0.0034, lty = 2)
     abline(h = 0.0021, lty = 2)
     abline(h = 0.0065, lty = 2)
    
    # Add x-axis and y-axis labels
    axis(1, at=c(1:6), dat.Ha$party, cex.axis=0.7)
    axis(2)
    
    # Add a frame
    box()

    dev.off() # close the current PDF file

```



