---
title: "Politbarometer Approval Ratings — Data Preparation"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

Reads the Politbarometer SAV files (2002–2021), extracts pre-election Skalometer ratings for
Spitzenkandidaten, and saves the result as `Data/politskalo.RData`. Run this once before
`Residual_Analysis.Rmd`. No 2025 wave is currently available.

```{r setup, message=FALSE, warning=FALSE}
library(haven)
```

# Variable notes

### 2021
V106: Baerbock (GR), KK
V110: Laschet (CDU), KK
V112: Lindner (FDP)
V116: Scholz (SPD), KK
V112: Wagenknecht (LI)
Missing: Weidel (AfD), Chrupalla (AfD), Wissler (LI), Bartsch (LI)
extra var: partly asked in 'wichtiger politiker' variable

### 2017
V104: Saklo Lindner (FDP)
V107: Skalo Merkel (CDU), KK
V109: Saklo Özdemir (GR)
V111: Skalo Schulz (SPD), KK
V114: Skalo Wagenknecht (LI)
missing: Bartsch (LI), Göring-Eckardt (GR), Weidel (AfD), Gauland (AfD).

### 2013
V114: Brüderle (FDP)
V119: Merkel (CDU), KK
V121: Gysi (LI)
V124: Steinbrück (SPD), KK
V126: Trittin (GR)
Missing: Göring-Eckardt (GR), Wagenknecht (LI), Bartsch (LI), Ernst (LI), van Aken (LI), Lay (LI), Gohlke (LI), Golze (LI).

### 2009 west
V105: Gysi (LI)
V108: Lafontaine (LI)
V109: Merkel (CDU), KK
V116: Steinmeier (SPD), KK
V117: Westerwelle (FDP)
Missing: Trittin (GR), Künast (GR)

### 2009 east
V105: Gysi (LI)
V108: Lafontaine (LI)
V109: Merkel (CDU), KK
V116: Steinmeier (SPD), KK
V117: Westerwelle (FDP)
Missing: Trittin (GR), Künast (GR)

### 2005 west
v143: Fischer (GR)
v144: Gysi (LI)
v147: Lafontaine (LI)
v148: Merkel (CDU)
v154: Schröder (SPD)
v158: Westerwelle (FDP)
Mssing: none
extra vars: preferred chancellor

### 2005 east
v143: Fischer (GR)
v144: Gysi (LI)
v147: Lafontaine (LI)
v148: Merkel (CDU)
v154: Schröder (SPD)
v158: Westerwelle (FDP)
Missing: none
extra vars: preferred chancellor

### 2002 west
v94: Fischer (GR)
v95: Gysi (LI) -> but we don't have him as SK, so they didn't have SK?
v103: Schröder (SPD), KK
v105: Stoiber (CDU), KK
v108: Westerwelle (FDP), KK
Missing: none
extra: here Künast (GR), v97 Merkel

### 2002 east
v94: Fischer (GR)
v95: Gysi (LI)
v103: Schröder (SPD), KK
v105: Stoiber (CDU), KK
v108: Westerwelle (FDP), KK
Missing: none
extra: here Künast (GR), v97 Merkel


# Config and helpers

The Skalometer runs from 1 (very negative) to 11 (very positive). We use the September wave in
each year, which is closest to the election. Where a politician does not appear in September
(the questionnaire rotates), we fall back to the single latest available pre-election month.

For years where the Politbarometer was fielded as separate east and west samples (2002, 2005,
2009), we combine both files so that the national mean is weighted by actual sample size rather
than giving equal weight to each half.

```{r pb_config}
pb.config <- list(
  `2021` = list(
    type = "combined",
    file = "../Data/politbarometer/ZA7856_v1-1-0_2021.sav",
    month.var = "V4", bundesland.var = "V6", election.month = 9,
    vars = c(
      "Baerbock, Annalena" = "V106",
      "Laschet, Armin"     = "V110",
      "Lindner, Chrisitan" = "V112",
      "Scholz, Olaf"       = "V116",
      "Wagenknecht, Sarah" = "V122"
    ),
    meta.vars = c(
      # Identifiers (no marital.status in 2021 — variable not collected that year)
      study.id          = "V1",   respondent.id     = "V2",
      survey.month      = "V4",   survey.week       = "V5",
      east.west         = "V7",
      municipality.size = "V8",   landline.mobile   = "V340",
      # Demographics
      gender            = "V324", age               = "V325",
      education         = "V326", state.exam        = "V328",
      vocational.train  = "V329", employment        = "V330",
      job.at.risk       = "V331", occupation        = "V332",
      hh.size           = "V333", hh.adults         = "V334",
      union.member      = "V335", religion          = "V336",
      church.attendance = "V337", party.id          = "V338",
      party.id.strength = "V339",
      # Weights
      weight.repr       = "V341", weight.total      = "V342"
    )
  ),
  `2017` = list(
    type = "combined",
    file = "../Data/politbarometer/ZA6988_v1-2-0_2017.sav",
    month.var = "V4", bundesland.var = "V7", election.month = 9,
    vars = c(
      "Lindner, Christian"  = "V104",
      "Merkel, Angela"      = "V107",
      "Özdemir, Cem"        = "V109",
      "Schulz, Martin"      = "V111",
      "Wagenknecht, Sarah"  = "V114"
    ),
    meta.vars = c(
      # Identifiers (V6 = landline/mobile in 2017; shifted: V8 = east.west, V9 = municipality)
      study.id          = "V1",   respondent.id     = "V2",
      survey.month      = "V4",   survey.week       = "V5",
      landline.mobile   = "V6",   east.west         = "V8",
      municipality.size = "V9",
      # Demographics
      gender            = "V317", age               = "V318",
      marital.status    = "V319", cohabit.partner   = "V320",
      education         = "V321", state.exam        = "V323",
      vocational.train  = "V324", employment        = "V325",
      job.at.risk       = "V326", occupation        = "V327",
      hh.size           = "V328", hh.adults         = "V329",
      union.member      = "V330", religion          = "V331",
      church.attendance = "V332", party.id          = "V333",
      party.id.strength = "V334",
      # Weights
      weight.repr       = "V335", weight.total      = "V336"
    )
  ),
  `2013` = list(
    type = "combined",
    file = "../Data/politbarometer/ZA5677_v1-1-0_2013.sav",
    month.var = "V4", bundesland.var = "V6", election.month = 9,
    vars = c(
      "Brüderle, Rainer" = "V114",
      "Merkel, Angela"   = "V119",
      "Gysi, Gregor"     = "V121",
      "Steinbrück, Peer" = "V124",
      "Trittin, Jürgen"  = "V126"
    ),
    meta.vars = c(
      # Identifiers
      study.id          = "V1",   respondent.id     = "V2",
      survey.month      = "V4",   survey.week       = "V5",
      east.west         = "V7",   municipality.size = "V8",
      # Demographics
      gender            = "V411", age               = "V412",
      marital.status    = "V413", cohabit.partner   = "V414",
      education         = "V430", vocational.train  = "V433",
      employment        = "V434", job.at.risk       = "V435",
      occupation        = "V436", hh.size           = "V437",
      hh.adults         = "V438", union.member      = "V439",
      religion          = "V440", church.attendance = "V441",
      party.id          = "V442", party.id.strength = "V443",
      # Weights
      weight.repr       = "V446", weight.total      = "V447"
    )
  ),
  `2009` = list(
    type = "split",
    file.west = "../Data/politbarometer/ZA5432_v1-2-0_2009_west.sav",
    file.east = "../Data/politbarometer/ZA5431_v1-2-0_2009_east.sav",
    month.var = "V4", bundesland.var = "V6", election.month = 9,
    vars = c(
      "Gysi, Gregor"       = "V105",
      "Lafontaine, Oskar"  = "V108",
      "Merkel, CDU"        = "V109",
      "Steinmeiner, SPD"   = "V116",
      "Westerwelle, Guido" = "V117"
    ),
    meta.vars = c(
      # Identifiers
      study.id          = "V1",   respondent.id     = "V2",
      survey.month      = "V4",   survey.week       = "V5",
      east.west         = ".src", municipality.size = "V8",
      # Demographics
      gender            = "V355", age               = "V356",
      marital.status    = "V357", cohabit.partner   = "V358",
      education         = "V374", state.exam        = "V376",
      vocational.train  = "V377", employment        = "V378",
      job.at.risk       = "V379", occupation        = "V380",
      hh.size           = "V381", hh.adults         = "V382",
      union.member      = "V383", religion          = "V384",
      church.attendance = "V385", party.id          = "V388",
      party.id.strength = "V389",
      # Weights
      weight.repr       = "V392", weight.total      = "V393"
    )
  ),
  `2005` = list(
    type = "split",
    file.west = "../Data/politbarometer/ZA4259_v1-1-0_2005_west.sav",
    file.east = "../Data/politbarometer/ZA4258_v1-1-0_2005_east.sav",
    month.var = "v3", bundesland.var = "v5", election.month = 9,
    vars = c(
      "Fischer, Joschka"   = "v143",
      "Gysi, Gregor"       = "v144",
      "Lafontaine, Oskar"  = "v147",
      "Merkel, CDU"        = "v148",
      "Schröder, SPD"      = "v154",
      "Westerwelle, Guido" = "v158"
    ),
    meta.vars = c(
      # Identifiers (no survey area; no landline/mobile)
      study.id          = "v1",   respondent.id     = "v2",
      survey.month      = "v3",   survey.week       = "v4",
      east.west         = ".src", municipality.size = "v6",
      # Demographics
      gender            = "v464", age               = "v465",
      marital.status    = "v466", cohabit.partner   = "v467",
      education         = "v483", state.exam        = "v484",
      vocational.train  = "v485",
      employment        = "v486", job.at.risk       = "v487",
      occupation        = "v488",
      hh.size           = "v489", hh.adults         = "v490",
      union.member      = "v494",
      religion          = "v495", church.attendance = "v496",
      party.id          = "v501", party.id.strength = "v502",
      # Weights
      weight.repr       = "v505", weight.total      = "v506"
    )
  ),
  `2002` = list(
    type = "split",
    file.west = "../Data/politbarometer/ZA3849_2002_west.sav",
    file.east = "../Data/politbarometer/ZA3850_2002_east.sav",
    month.var = "v3", bundesland.var = "v5", election.month = 9,
    vars = c(
      "Fischer, Joschka"   = "v94",
      "Schröder, SPD"      = "v103",
      "Stoiber, CDU"       = "v105",
      "Westerwelle, Guido" = "v108"
    ),
    meta.vars = c(
      # Identifiers (no version; v4 = survey number not week; no landline/mobile)
      study.id          = "v1",   respondent.id     = "v2",
      survey.month      = "v3",   survey.number     = "v4",
      east.west         = ".src", municipality.size = "v6",
      # Demographics (no job.at.risk, no union.member in 2002)
      gender            = "v268", age               = "v269",
      marital.status    = "v270", cohabit.partner   = "v271",
      education         = "v287", state.exam        = "v288",
      vocational.train  = "v289",
      employment        = "v290", occupation        = "v292",
      hh.size           = "v293", hh.adults         = "v294",
      religion          = "v299", church.attendance = "v300",
      party.id          = "v303", party.id.strength = "v304",
      # Weights
      weight.repr       = "v308", weight.total      = "v309"
    )
  )
)

# Uniform 1-16 state numbers keyed by label (Berlin West+Ost both → 11).
bl.num.map <- c(
  "Schleswig-Holstein"     =  1, "Hamburg"               =  2,
  "Niedersachsen"          =  3, "Bremen"                =  4,
  "NRW"                    =  5, "Hessen"                =  6,
  "Rheinland-Pfalz"        =  7, "Baden-Württemberg"     =  8,
  "Bayern"                 =  9, "Saarland"              = 10,
  "Berlin"                 = 11, "Brandenburg"           = 12,
  "Mecklenburg-Vorpommern" = 13, "Sachsen"               = 14,
  "Sachsen-Anhalt"         = 15, "Thüringen"             = 16
)

# Politbarometer codes 1-17, splitting Berlin into West (11) and Ost (12).
# Both merged to "Berlin" to align with the 16-state BWL system.
bl.label.map <- c(
  # Standard codes 1-17 (all years)
  `1`="Schleswig-Holstein", `2`="Hamburg",          `3`="Niedersachsen",
  `4`="Bremen",             `5`="NRW",               `6`="Hessen",
  `7`="Rheinland-Pfalz",   `8`="Baden-Württemberg", `9`="Bayern",
  `10`="Saarland",         `11`="Berlin",           `12`="Berlin",
  `13`="Brandenburg",      `14`="Mecklenburg-Vorpommern",
  `15`="Sachsen",          `16`="Sachsen-Anhalt",   `17`="Thüringen",
  # 2002 east file uses codes 21-26 instead of 12-17
  `21`="Berlin",           `22`="Brandenburg",
  `23`="Mecklenburg-Vorpommern", `24`="Sachsen",
  `25`="Sachsen-Anhalt",   `26`="Thüringen"
)

# Returns one row per respondent x politician from the election-month wave.
# Columns: candidates, rating (1-11), bundesland, bundesland.label, plus all meta.vars.
extract.raw.ratings <- function(dat, vars, month.var, election.month, bundesland.var,
                                meta.vars = NULL) {
  month.num <- as.numeric(dat[[month.var]])
  in.wave   <- !is.na(month.num) & month.num == election.month
  if (!any(in.wave)) {
    pre.months <- month.num[month.num < election.month & !is.na(month.num)]
    if (length(pre.months) == 0) return(NULL)
    m <- max(pre.months)
    cat("  Note: election month", election.month, "empty; using month", m, "\n")
    in.wave <- !is.na(month.num) & month.num == m
  }
  bl.raw <- as.numeric(dat[[bundesland.var]])
  keep   <- in.wave
  rows <- lapply(names(vars), function(nm) {
    v      <- vars[[nm]]
    rating <- suppressWarnings(as.numeric(dat[[v]][keep]))
    rating[!is.na(rating) & (rating < 1 | rating > 11)] <- NA
    if (!any(keep)) return(NULL)
    df <- data.frame(
      candidates       = nm,
      rating           = rating,
      bundesland       = bl.raw[keep],
      bundesland.label = bl.label.map[as.character(bl.raw[keep])],
      state            = bl.num.map[bl.label.map[as.character(bl.raw[keep])]],
      stringsAsFactors = FALSE
    )
    if (!is.null(meta.vars)) {
      for (std_nm in names(meta.vars)) {
        v_meta <- meta.vars[[std_nm]]
        df[[std_nm]] <- if (v_meta %in% names(dat))
          suppressWarnings(as.numeric(dat[[v_meta]][keep]))
        else NA_real_
      }
    }
    df
  })
  rows <- Filter(Negate(is.null), rows)
  if (length(rows) == 0) return(NULL)
  do.call(rbind, rows)
}

# Returns individual-level long data for all years.
# Columns: candidates, year, rating, bundesland, bundesland.label, plus meta columns.
build.approval.long <- function(cfg) {
  year.rows <- lapply(names(cfg), function(yr) {
    c. <- cfg[[yr]]
    dat <- if (c.$type == "combined") {
      zap_labels(read_sav(c.$file))
    } else {
      w <- zap_labels(read_sav(c.$file.west)); w$.src <- 1L
      e <- zap_labels(read_sav(c.$file.east)); e$.src <- 2L
      dplyr::bind_rows(w, e)
    }
    raw <- extract.raw.ratings(dat, c.$vars, c.$month.var, c.$election.month,
                               c.$bundesland.var, c.$meta.vars)
    if (is.null(raw)) return(NULL)
    raw$year <- as.numeric(yr)
    raw
  })
  lst <- Filter(Negate(is.null), year.rows)
  all_cols <- unique(unlist(lapply(lst, names)))
  do.call(rbind, lapply(lst, function(df) {
    df[setdiff(all_cols, names(df))] <- NA
    df[all_cols]
  }))
}
```

# Build and save

```{r build_save}
politskalo <- build.approval.long(pb.config)
# Mobile sampling introduced in 2017. Before that: landline only (code 2 = Festnetzstichprobe).
politskalo$landline.mobile[is.na(politskalo$landline.mobile)] <- 2L

# Harmonise religion to the 6-category 2005+ scheme.
# 2002 has only 5 categories (no Jude/Jüdin): 4=andere, 5=keiner.
# 2005+ scheme:                                4=Jude/Jüdin, 5=andere, 6=keiner.
# Shift 2002 codes 4→5 and 5→6 simultaneously (ifelse evaluates in one pass).
r2002 <- !is.na(politskalo$religion) & politskalo$year == 2002
politskalo$religion <- ifelse(r2002 & politskalo$religion == 4L, 5L,
                       ifelse(r2002 & politskalo$religion == 5L, 6L,
                              politskalo$religion))
# After recoding: code 4 (jüdisch) has no 2002 observations — correct,
# since 2002 absorbed that group into "andere" with no separate category.

save(politskalo, file = "../Data/politskalo.RData")
cat("Saved politskalo to Data/politskalo.RData\n")
cat("Rows:", nrow(politskalo), "\n")
cat("Years covered:", paste(sort(unique(politskalo$year)), collapse = ", "), "\n")
cat("Candidates:", paste(sort(unique(politskalo$candidates)), collapse = ", "), "\n")
```
